---
title: "HCC_proteome"
output: 
  html_document: default
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###CPTAC proteome

```{r}
library(dplyr)
prodata <- data.table::fread("./S049_Liver_Cancer_Gao2019/Liver_Cancer_Proteome_CDAP_Protein_Report.r1/Zhou_Liver_Cancer_Proteome.tmt11.tsv", header = T)%>% as.data.frame()
rt <- prodata[!duplicated(prodata$Gene),]
rownames(rt)<-rt[,1]
rt<-rt[,-1]
rt=as.matrix(rt)

selectCol=seq(2,ncol(rt),2) 
rt=rt[,selectCol]

hcc_data <- as.data.frame(rt)
openxlsx::write.xlsx(hcc_data,file="hcc_prodata2.xlsx", rowNames = T, colName = T)
```


###TCGA_data
```{r}
library(limma)
library(tibble)
HCC_data <- data.table::fread("C:/Users/64684/Documents/R/LIHC/TCGA_LIHCdata/TCGA_HCC_FPKM.txt", header = T)
Normal_data <- data.table::fread("C:/Users/64684/Documents/R/LIHC/TCGA_LIHCdata/TCGA_Normal_FPKM.txt", header = T)

HCC_matrix <- merge(HCC_data, Normal_data, by="GeneSymbol")
HCC_matrix <- column_to_rownames(HCC_matrix, var = "GeneSymbol")

groups <- c(rep("TT",369), rep("NAT",50))

design<- model.matrix(~0+factor(groups))
colnames(design)<- levels(factor(groups))

fit <- lmFit(HCC_matrix, design)

cont.matrix <- makeContrasts('TT-NAT', levels = design) 
fit2 <- contrasts.fit(fit, cont.matrix)
fit2 <- eBayes(fit2)

output <- topTable(fit2, coef = 1, number = Inf, adjust.method = 'BH')
write.csv(output, file = "LIHC_DEGs.csv",row.names = T)


```


##HCC_proteome
###Tumor tissue vs peritumor tissue
#### Lollipop chart
```{r}
library(dplyr)
library(ggplot2)
library(hrbrthemes)
library(ggpubr)
no.data <- openxlsx::read.xlsx("./2022MBR/Proteome_MBR.xlsx", sheet = 3)
lollipop <- no.data[,c(4:6)]

g <- ggplot(lollipop) +
  geom_segment(aes(x=SampleID, xend=SampleID, y=Peritumor, yend=Tumor),
               color="grey") +
  geom_point(aes(x=SampleID, y=Peritumor), color="#2c83b7", size=2 ) +
  geom_point(aes(x=SampleID, y=Tumor), color="#c2395a", size=2 ) +
  theme(legend.position = "none") +
  theme_classic() + ylim(5000,8000)+
  xlab("Paired tumor and peritumor samples (n = 39)") +
  ylab("Number of protein identifications")

plotdata <- no.data[,c(1:3)]

p <- ggplot(plotdata, aes(x = SampleID, y = No.)) + 
  geom_point(aes(color = Groups), size=2) + 
  geom_smooth(aes(color = Groups))+
  theme_classic() + ylim(5000,8000)+
  theme(legend.position = "none") +
  xlab("Paired tumor and peritumor samples (n = 39)") +
  ylab("Number of protein identifications")

b <- ggplot(plotdata, aes(x = Groups, y = No.)) + 
  geom_jitter(aes(color = Groups), size=1, width = 0.3)+
  geom_boxplot(aes(color = Groups), size=1) + 
  scale_color_manual(values=c(Peritumor = "#2c83b7", Tumor = "#c2395a"))+
  stat_compare_means()+
  theme_classic() + ylim(5000,8000)+
  theme(legend.position = "none") +
  xlab("") +
  ylab("Number of protein identifications")

```

#### Venn graph
```{r}
library(VennDiagram)
library(grDevices)
library(dplyr)
filedata <- openxlsx::read.xlsx("./2022MBR/Venn/Venngene_list.xlsx", sheet = 1)%>%as.data.frame()
p <- venn.diagram(
     list(CPTAC=filedata$CPTAC, This_Study=filedata$This.Study), 
     na = "remove", filename = NULL,  col = "black", lty = 1, lwd = 3,
     fill = c("cornflowerblue", "darkorchid1"), alpha = 0.50, cex = 2.0,
     fontfamily = "serif", fontface = "bold",
     cat.col = c("cornflowerblue", "darkorchid4"), cat.cex = 1.8,
     cat.fontface = "bold",  cat.fontfamily = "serif")

pdf(file="2quadruple_Venn.pdf")
grid.draw(p)
dev.off()

pt <- venn.diagram(
     list(Tumor=filedata$Tumor, Peritumor=filedata$Peritumor), 
     na = "remove", filename = NULL,  lty = 1, lwd = 3,
     col = c('#c2395a', '#2c83b7'), alpha = 0.50, cex = 2.0,
     fontfamily = "serif", fontface = "bold",
     cat.col = c('#c2395a', '#2c83b7'), cat.cex = 1.8,
     cat.fontface = "bold",  cat.fontfamily = "serif")

pdf(file="2PTvsTT_Venn.pdf")
grid.draw(pt)
dev.off()
```

#### Metabolites Count Pieplot
```{r}
library(ggplot2)
library(RColorBrewer)
metabolite_data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolites2023.xlsx", sheet = 1, rowNames = F)

count <- as.data.frame(table(metabolite_data$Super_Class))
write.csv(count, file = "Metablites_class.csv", row.names = F)

count <- read.csv("./2022ProteomeMetabolism/Metablites_class.csv")
count <- count[order(count$Var1, decreasing = TRUE),]

p <- ggplot(data = count, mapping = aes(x = '', y = Freq, fill = Var1)) +
  geom_bar(stat = 'identity', position = 'stack', width = 1)+
  coord_polar(theta = 'y')

label_value <- paste('(', round(count$Freq/sum(count$Freq) * 100, 2), '%)', 
                     sep = '')
label_value

label <- paste(count$Var1, label_value, sep = '')
##blank theme
blank_theme <- theme_minimal()+
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.border = element_blank(),
  panel.grid=element_blank(),
  axis.ticks = element_blank(),
  plot.title=element_text(size=5, face="bold")
  )
g <- p + scale_fill_brewer(palette = "Paired")+
     labs(x = '', y = '', title = '') + blank_theme +
     theme(legend.position = "none") + 
     theme(axis.text.x=element_blank()) +
     geom_text(aes(y = count$Freq/5 +
                     c(0,cumsum(count$Freq)[-length(count$Freq)]),
                   x = sum(count$Freq)/600, label = label)) 

```

##### Venn graph of metabolites
```{r}
library(VennDiagram)
library(grDevices)
library(dplyr)
filedata <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolites2023.xlsx", sheet = 4)%>%as.data.frame()
p <- venn.diagram(
     list(Negative=filedata$Negative, Positive=filedata$Positive), 
      na = "remove", filename = NULL,  lty = 1, lwd = 3,
     col = c('#c2395a', '#2c83b7'), alpha = 0.50, cex = 2.0,
     fontfamily = "serif", fontface = "bold",
     cat.col = c('#c2395a', '#2c83b7'), cat.cex = 1.8,
     cat.fontface = "bold",  cat.fontfamily = "serif")

pdf(file="2quadruple_metabolites.pdf")
grid.draw(p)
dev.off()

```

####Lipid Count Pieplot
```{r}
library(ggplot2)
library(RColorBrewer)
lipid_data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Lipid2023.xlsx", sheet = 1, rowNames = F)

count <- as.data.frame(table(lipid_data$Category))
write.csv(count, file = "Lipid_class.csv", row.names = F)

count <- read.csv("./2022ProteomeMetabolism/Lipid_class.csv")
count <- count[order(count$Var1, decreasing = TRUE),]

p <- ggplot(data = count, mapping = aes(x = '', y = Freq, fill = Var1)) +
  geom_bar(stat = 'identity', position = 'stack', width = 1)+
  coord_polar(theta = 'y')

label_value <- paste('(', round(count$Freq/sum(count$Freq) * 100, 2), '%)', 
                     sep = '')
label_value

label <- paste(count$Var1, label_value, sep = '')
##blank theme
blank_theme <- theme_minimal()+
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.border = element_blank(),
  panel.grid=element_blank(),
  axis.ticks = element_blank(),
  plot.title=element_text(size=5, face="bold")
  )
g <- p + scale_fill_brewer(palette = "Set1")+
     labs(x = '', y = '', title = '') + blank_theme +
     theme(legend.position = "none") + 
     theme(axis.text.x=element_blank()) +
     geom_text(aes(y = count$Freq/10 +
                     c(0,cumsum(count$Freq)[-length(count$Freq)]),
                   x = sum(count$Freq)/2000, label = label)) 

```

##### Venn graph of lipids
```{r}
library(VennDiagram)
library(grDevices)
library(dplyr)
filedata <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Lipid2023.xlsx", sheet = 4)%>%as.data.frame()
p <- venn.diagram(
     list(Negative=filedata$Negative, Positive=filedata$Positive), 
      na = "remove", filename = NULL,  lty = 1, lwd = 3,
     col = c('#c2395a', '#2c83b7'), alpha = 0.50, cex = 2.0,
     fontfamily = "serif", fontface = "bold",
     cat.col = c('#c2395a', '#2c83b7'), cat.cex = 1.8,
     cat.fontface = "bold",  cat.fontfamily = "serif")

pdf(file="2quadruple_lipids.pdf")
grid.draw(p)
dev.off()

```

####Lipid Count Boxplot
```{r}
library(ggplot2)
library(ggpubr)
data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Lipid_merge_Rawdata.xlsx", sheet = 2)

p <- ggplot(data, aes(x=Segments, y=No., color=Groups))+
  geom_boxplot()+
  scale_color_manual(values=c(TT="#c2395a", PT="#2c83b7"))+
  theme_classic()+
  theme(legend.position = "right", 
        legend.key.height = unit(0.3,'cm'),
        legend.key.width = unit(0.3,'cm'))+
  xlab("Segments") + ylab("Numbers of Detected Lipids")


```

####Proteome_PCA
```{r}
library(FactoMineR)
library(ggplot2)
library(dplyr)
prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1,rowNames = T)

plotdata <- as.data.frame(t(log2(prodata)))

prodata.pca <- PCA(plotdata, ncp = 2, scale.unit = TRUE, graph = FALSE)

plot(prodata.pca)

pca_data <- data.frame(prodata.pca$ind$coord[,1:2])

pca_eig1 <- round(prodata.pca$eig[1,2], 2)
pca_eig2 <- round(prodata.pca$eig[2,2], 2)

groups <- as.data.frame(c(rep("NAT",39), rep("TT",39)))
names(groups) <- "Groups"

pca_data <- cbind(pca_data, groups)
pca_data$SampleID <- rownames(pca_data)

p <- ggplot(data = pca_data, aes(x = Dim.1, y = Dim.2)) +
  geom_point(aes(color = Groups), size = 1) +
  scale_color_manual(values = c('#2c83b7', '#c2395a')) + 
  theme_bw() +
  theme(legend.position = "none",
        panel.grid=element_blank(),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14)) +
  labs(x =  paste('PCA1:', pca_eig1, '%'), 
       y = paste('PCA2:', pca_eig2, '%'), color = '')

g <- p + stat_ellipse(aes(fill = Groups), geom = 'polygon', level = 0.95, 
                 alpha = 0.1, show.legend = FALSE) +
    scale_fill_manual(values = c('#2c83b7', '#c2395a'))


```

#### Metabolomics OPLS-DA
```{r}
library(ropls)
Metabolite_data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolites2023.xlsx", sheet = 1,rowNames = T)

group <- openxlsx::read.xlsx("./2022proteome/Clinicaldata2.xlsx", sheet = 2,rowNames = F)
##PCA analysis
Metabolite.pca <- opls(t(log2(Metabolite_data[,-c(1:12)])))

plot(Metabolite.pca,
     typeVc = "x-score",
     parAsColFcVn = group$Groups,
     parPaletteVc = c("#2c83b7", "#c2395a"))

##PLS-DA
pdf("Metabolite_plsda.pdf")
lipid.plsda <- opls(t(log2(Metabolite_data)), group$Groups)
dev.off()

plot(Metabolite.pca,
     typeVc = "x-score",
     parAsColFcVn = group$Groups,
     parPaletteVc = c("#2c83b7", "#c2395a"))

## OPLS-DA
pdf("Metabolite_oplsda.pdf")
Metabolite.oplsda <- opls(t(log2(Metabolite_data[,-c(1:12)])), group$Groups, predI = 1, orthoI = NA)
dev.off()

plot(Metabolite.oplsda,
     typeVc = "x-score",
     parAsColFcVn = group$Groups,
     parPaletteVc = c("#2c83b7", "#c2395a"))
```

#####ggplot2
```{r}
library(ggplot2)
library(ggsci)
library(tidyverse)

sample.score = Metabolite.oplsda@scoreMN %>%  
  as.data.frame() %>%
  mutate(group = group$Groups,
         o1 = Metabolite.oplsda@orthoScoreMN[,1]) 
head(sample.score)

p <- ggplot(sample.score, aes(p1, o1, color = group)) +
  geom_hline(yintercept = 0, linetype = 'dashed', size = 0.5) + 
  geom_vline(xintercept = 0, linetype = 'dashed', size = 0.5) +
  geom_point() +
  #geom_point(aes(-10,-10), color = 'white') +
  labs(x = 'OPLS-DA:t1(6.0%)',y = 'OPLS-DA:to1') +
  stat_ellipse(level = 0.95, linetype = 'solid', 
               size = 1, show.legend = FALSE) + 
  scale_color_manual(values = c('#2c83b7','#c2395a')) +
  theme_bw() +
  theme(legend.position = c(0.92,0.92),
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16))
```

#### Lipidomics OPLS-DA
```{r}
library(ropls)
Lipid_data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Lipid2023.xlsx", sheet = 1,rowNames = T)

group <- openxlsx::read.xlsx("./2022proteome/Clinicaldata2.xlsx", sheet = 2,rowNames = F)

## OPLS-DA
pdf("Lipid_data_oplsda.pdf")
Lipid.oplsda <- opls(t(log2(Lipid_data[,-1])), group$Groups, predI = 1, orthoI = NA)
dev.off()

plot(Lipid.oplsda,
     typeVc = "x-score",
     parAsColFcVn = group$Groups,
     parPaletteVc = c("#2c83b7", "#c2395a"))
```
#####ggplot2 
```{r}
library(ggplot2)
library(ggsci)
library(tidyverse)
sample.score = Lipid.oplsda@scoreMN %>% 
  as.data.frame() %>%
  mutate(group = group$Groups,
         o1 = Lipid.oplsda@orthoScoreMN[,1]) 
head(sample.score)

p <- ggplot(sample.score, aes(p1, o1, color = group)) +
  geom_hline(yintercept = 0, linetype = 'dashed', size = 0.5) + 
  geom_vline(xintercept = 0, linetype = 'dashed', size = 0.5) +
  geom_point() +
  #geom_point(aes(-10,-10), color = 'white') +
  labs(x = 'OPLS-DA:t1(6.0%)',y = 'OPLS-DA:to1') +
  stat_ellipse(level = 0.95, linetype = 'solid', 
               size = 1, show.legend = FALSE) + 
  scale_color_manual(values = c('#2c83b7','#c2395a')) +
  theme_bw() +
  theme(legend.position = c(0.92,0.92),
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16))
```

##### Indentifing the DALs and DAMs
```{r}
lipid.vip <- as.data.frame(getVipVn(Metabolite.oplsda))
write.csv(lipid.vip, file = "Metabolite_VIP.csv", row.names = T)

```

#### Liver specific proteins
```{r}
library(tibble)
library(reshape2)
library(ggplot2)
library(ggpubr)
prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1,rowNames = F)

proteinlist <- openxlsx::read.xlsx("./2022MBR_filter/Liver_specific_genes.xlsx", sheet = 1,rowNames = F)
proteinlist <- as.data.frame(proteinlist[,1])
names(proteinlist) <- "Name"
specific_protein_exp <- merge(proteinlist, prodata, by="Name")
specific_protein_exp <- column_to_rownames(specific_protein_exp, var = "Name")
 
plotdata <- as.data.frame(t(log2(specific_protein_exp+1)))
groups <- as.data.frame(c(rep("NAT",39), rep("TT",39))) 
names(groups) <- "Groups"

plotdata <- cbind(groups, plotdata)

data <- melt(plotdata, id.vars = "Groups")

p <- ggplot(data, aes(x=Groups, y=value, color=Groups))+geom_boxplot()+
  scale_color_manual(values=c(NAT = "#2c83b7", TT = "#c2395a"))+
  stat_compare_means()+
  theme_classic()+ 
  ylab("Log2(Normalized expression value)")+
  xlab("Liver specific proteins")+
  theme(legend.position = "none",
        panel.grid=element_blank(),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))
```

#### metabolite-interacting proteins (MIPros)
```{r}
library(tibble)
library(reshape2)
library(ggplot2)
library(ggpubr)
prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1,rowNames = F)

proteinlist <- openxlsx::read.xlsx("./2022MBR_filter/MIPros_list.xlsx", sheet = 2,rowNames = F)

MIPros_exp <- merge(proteinlist, prodata, by="Name")
MIPros_exp <- column_to_rownames(MIPros_exp, var = "Name")
 
plotdata <- as.data.frame(t(log2(MIPros_exp+1)))
groups <- as.data.frame(c(rep("NAT",39), rep("TT",39))) 
names(groups) <- "Groups"

plotdata <- cbind(groups, plotdata)

data <- melt(plotdata, id.vars = "Groups")

p <- ggplot(data, aes(x=Groups, y=value, color=Groups))+
  geom_boxplot(outlier.size = 0.5,alpha=0.5)+
  scale_color_manual(values=c(NAT = "#2c83b7", TT = "#c2395a"))+
  stat_compare_means()+
  theme_classic()+ 
  ylab("Log2(Normalized expression value)")+
  xlab("Proteins assigned to metabolic pathways")+
  theme(legend.position = "none",
        panel.grid=element_blank(),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))
```

#### DEP in Tumor vs NAT
```{r}
library(dplyr)
library(limma)
prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1,rowNames = T)

groups <- c(rep("NAT",39), rep("TT",39))

design<- model.matrix(~0+factor(groups))
colnames(design)<- levels(factor(groups))
##voom transform
v <- voom(prodata, design, plot=TRUE)

fit <- lmFit(v, design)

cont.matrix <- makeContrasts('TT-NAT', levels = design) 
fit2 <- contrasts.fit(fit, cont.matrix)
fit2 <- eBayes(fit2)

output <- topTable(fit2, coef = 1, number = Inf, adjust.method = 'BH')
write.csv(output, file = "Proteome_DEGs_limmavoom2.csv",row.names = T)

```

>metabolites

```{r}
library(dplyr)
library(limma)
lipid_data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolites2023.xlsx", sheet = 1,rowNames = T)%>% as.data.frame()

group <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 3, rowNames = T)
design<- model.matrix(~0+factor(group$Groups))
colnames(design)<- levels(factor(group$Groups))  

v <- voom(lipid_data[,-1], design, plot=TRUE)
v <- voom(lipid_data[,-c(1:12)], design, plot=TRUE)

fit <- lmFit(v, design)

cont.matrix <- makeContrasts('TT-NAT', levels = design) 
fit2 <- contrasts.fit(fit, cont.matrix)
fit2 <- eBayes(fit2)

output <- topTable(fit2, coef = 1, number = Inf, adjust.method = 'BH')
write.csv(output, file = "Metabolite_limmavoom.csv",row.names = T)
```


#### Volcano plot
```{r}
library(ggplot2)
library(ggrepel)
names(output) <- c('log2FC', 'AveExpr', 't', 'Pvalue', 'BH-Pvalue', 'B')
yanse <- ifelse(output$`BH-Pvalue`<0.01 & abs(output$log2FC)>= 1, ifelse(output$log2FC> 1,'red','gray'),'green')
p <- ggplot(output, aes(output$log2FC, -log10(output$`BH-Pvalue`), col=yanse))
g <- p+geom_point(size=0.5)+theme_bw()+ 
  scale_color_manual(values = c('#2c83b7','grey60','#c2395a'))+ 
  labs(x="log2 (Fold Change)",y="-log10 (BH-Pvalue)") +
  geom_hline(yintercept = -log10(0.05), lty=4,col="grey",lwd=0.6) +
  geom_vline(xintercept = c(-1, 1), lty=4,col="grey",lwd=0.6) +
  xlim(-5,5)+
  theme(legend.position = "none",
        panel.grid=element_blank(),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))
##label
label <- ifelse(-log10(output$`BH-Pvalue`) > 10 & abs(output$log2FC) >= 2, rownames(output),"")

l <- g+geom_text_repel(data=output, aes(x = output$log2FC, 
                                   y = -log10(output$`BH-Pvalue`), 
                                   label = label),
                  size = 3,box.padding = unit(0.2, "lines"),
                  segment.color = "black", 
                  show.legend = FALSE)

```

####heatmap of DEPs in TT vs PT
```{r}
library(ComplexHeatmap)
library(tibble)
library(dplyr)
DEP_list <- openxlsx::read.xlsx("./Proteome_DEGs_limmavoom.xlsx", sheet = 3)
DEP_list <- DEP_list[,1]%>%as.data.frame()
names(DEP_list) <- "Name"

prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1,rowNames = F)

#prodata <- rownames_to_column(prodata, var = "Name")

DEP_exp <- merge(DEP_list, prodata, by="Name", sort = F)
DEP_exp <- column_to_rownames(DEP_exp, var = "Name")

plotdata <- as.data.frame(t(scale(log2(t(DEP_exp)))))

groups <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 3)

mark_gene <- openxlsx::read.xlsx("./Proteome_DEGs_limmavoom.xlsx", sheet = 4)%>%as.data.frame()

row_anno <-  rowAnnotation(link = anno_mark(at = which(rownames(plotdata) 
                                                       %in% mark_gene$Name), 
                                       labels = mark_gene$Name,
                           labels_gp = gpar(fontsize = 6)))

ht = HeatmapAnnotation(Cluster = groups$Groups,
                       show_legend = F,
                       col = list(Cluster = c("NAT" = "#2c83b7",
                                               "TT"   = "#c2395a")))
col_fun = circlize::colorRamp2(c(-4, -2, 0, 2, 4),
                               c("#2c83b7","deepskyblue1", "white",
                                 "red", "#c2395a"))

p <- Heatmap(plotdata, 
             col = col_fun,
             top_annotation = ht,
             show_row_names = F,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = F,
             cluster_columns = F,
             cluster_rows = F,
             heatmap_legend_param = list(title = "Zscore"))
p+row_anno
```

>metabolites

```{r}
library(ComplexHeatmap)
library(tibble)
metabolites_exp <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolites2.xlsx", sheet = 2)
metabolites_list <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolites_limmavoom.xlsx", sheet = 2)

metabolites_list <- as.data.frame(metabolites_list[,1])
names(metabolites_list) <- "Name"

DEM_exp <- merge(metabolites_list, metabolites_exp, sort = F, by="Name")
DEM_exp <- column_to_rownames(DEM_exp, var = "Name")

plotdata <- as.data.frame(t(scale(t(log2(DEM_exp+1)))))

groups <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 3)

ht = HeatmapAnnotation(Cluster = groups$Groups,
                        show_legend = F,
                        col = list(Cluster = c("NAT" = "#2c83b7",
                                               "TT" = "#c2395a")))

col_fun = circlize::colorRamp2(c(-4, -2, 0, 2, 4),
                               c("#2c83b7","deepskyblue1", "white",
                                 "red", "#c2395a"))
marker <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolites_limmavoom.xlsx", sheet = 3)%>%as.data.frame()

row_anno <-  rowAnnotation(link = anno_mark(at = which(rownames(plotdata) 
                                                       %in% marker$Name), 
                                       labels = marker$Name,
                           labels_gp = gpar(fontsize = 6)))

p <- Heatmap(plotdata, 
             col = col_fun,
             top_annotation = ht, 
             show_row_names = F,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = F,
             cluster_columns = F,
             cluster_rows = F,
             row_names_gp = gpar(fontsize = 6),
             heatmap_legend_param = list(title = "Zscore", 
                                         direction = "horizontal"))

p+row_anno
```

>lipids

```{r}
library(ComplexHeatmap)
library(tibble)
lipid_exp <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Lipid2023.xlsx", sheet = 1)
lipid_list <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Lipids_DAL_VIP.xlsx", sheet = 2)

lipid_list <- as.data.frame(lipid_list[,1])
names(lipid_list) <- "Name"

DEM_exp <- merge(lipid_list, lipid_exp, sort = F, by="Name")
DEM_exp <- column_to_rownames(DEM_exp, var = "Name")

plotdata <- as.data.frame(t(scale(t(log2(DEM_exp[,-1]+1)))))

groups <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 3)

ht = HeatmapAnnotation(Cluster = groups$Groups,
                        show_legend = F,
                        col = list(Cluster = c("NAT" = "#2c83b7",
                                               "TT" = "#c2395a")))

col_fun = circlize::colorRamp2(c(-4, -2, 0, 2, 4),
                               c("#2c83b7","deepskyblue1", "white",
                                 "red", "#c2395a"))
marker <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Lipids_DAL_VIP.xlsx", sheet = 3)%>%as.data.frame()

row_anno <-  rowAnnotation(link = anno_mark(at = which(rownames(plotdata) 
                                                       %in% marker$Name), 
                                       labels = marker$Name,
                           labels_gp = gpar(fontsize = 6)))

p <- Heatmap(plotdata, 
             col = col_fun,
             top_annotation = ht, 
             show_row_names = F,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = F,
             cluster_columns = F,
             cluster_rows = F,
             row_names_gp = gpar(fontsize = 6),
             heatmap_legend_param = list(title = "Zscore", 
                                         direction = "horizontal"))

p+row_anno
```

### Metascale anotation
```{r}
library(ggplot2)

Tumor.data <- openxlsx::read.xlsx("./2022proteome/TTvsNAT-metascape.xlsx", sheet = 1)

tt <- ggplot(Tumor.data, aes(y = reorder(factor(Description), `-Log(q-value)`),
                             x = `-Log(q-value)`)) + 
  geom_bar(stat = 'identity', fill = '#c2395a', width = 0.7)+
  xlab("")+ylab("-log10(q-value)")+
  theme_classic() +
  theme(legend.position ="",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))

NAT.data <- openxlsx::read.xlsx("./2022proteome/TTvsNAT-metascape.xlsx", sheet = 2)

nat <- ggplot(NAT.data, aes(x = reorder(factor(Description), `-Log(q-value)`),
                            y = `-Log(q-value)`)) + 
     geom_bar(stat = 'identity', fill = "#2c83b7", width = 0.7) +
     xlab("")+ylab("-log10(p-value)")+
     coord_flip()+
     theme_classic() +
     theme(legend.position ="",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))

c2 <- ggplot(cluster2, aes(x = reorder(factor(Description), LogP), y = LogP)) + 
  geom_bar(stat = 'identity', fill = '#0f4392', width = 0.7)+
  xlab("")+ylab("-log10(p-value)")+
  coord_flip()+
  theme_classic() +
  theme(legend.position ="",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))

```

#### Bubble plot of MSEA in TTvsNAT
```{r}
library(ggplot2)

nat.data <- read.csv("./2022ProteomeMetabolism/02NAT_msea_ora_result.csv", header = T)

plotdata <- nat.data[c(1:13),c(1,8,9)]

p <- ggplot(plotdata,aes(x = LogP, 
              y = EnrichmentRatio, 
              size = EnrichmentRatio,
              colour=LogP)) +
     geom_point(shape = 16) +               
     labs(x = "LogP", y = "EnrichmentRatio")+          
     scale_colour_gradient2(name="-log10(P-value)",
     low = "#ffbd39",
     mid = "#e61c5d",
     high = "#930077",
     midpoint = 2)+
     theme_bw()          


tt.data <- read.csv("./2022ProteomeMetabolism/01TT_msea_ora_result.csv", header = T)

plotdata <- tt.data[c(1:10),c(1,8,9)]

```


### Kruskal-Wallis test in NAT
```{r}
library(dplyr)

prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1, rowNames = T)%>%as.data.frame()
kw_data <- as.data.frame(t(prodata[,c(1:39)]))

group <- openxlsx::read.xlsx("./2022proteome/Clinicaldata2.xlsx", sheet = 4, rowNames = T)%>%as.data.frame()

kw_data <- cbind(group[6], kw_data)

data1 <- kruskal.test(as.matrix(kw_data[2]) ~Segments, data = kw_data)

pval=c()

for (i in 2:ncol(kw_data)) {
  p = kruskal.test(as.matrix(kw_data[i]) ~Segments, data = kw_data)$p.value
  pval=c(pval,p)
  
}

table_out <- as.data.frame(cbind(row.names(prodata),pval))

write.csv(table_out, file = "NAT-KW-test_in_proteome.csv")

```

#### NAT KW-DEP heatmap
```{r}
library(ComplexHeatmap)
library(tibble)
prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1, rowNames = F)

NAT_DEP_list <- openxlsx::read.xlsx("./2022proteome/NAT-KW-test_in_proteome.xlsx", sheet = 2, rowNames = F)

NAT_DEP_exp <- merge(NAT_DEP_list[1], prodata[,1:40], by="Name")
NAT_DEP_exp <- column_to_rownames(NAT_DEP_exp, var = "Name")

plotdata <- as.data.frame(t(scale(t(log2(NAT_DEP_exp+1)))))

groups <- openxlsx::read.xlsx("./2022proteome/Clinicaldata2.xlsx", sheet = 4)

ht = HeatmapAnnotation(Segments = groups$Segments,
                       show_legend = T,
                       col = list(Segments = c("S1" = "#48466d",
                                               "S2"   = "#3d84a8",
                                               "S3"   = "#46cdcf",
                                               "S4"   = "#abedd8",
                                               "S5"   = "#f9ed69",
                                               "S6"   = "#f08a5d",
                                               "S7"   = "#b83b5e",
                                               "S8"   = "#6a2c70")))

col_fun = circlize::colorRamp2(c(-4, -2, 0, 2, 4),
                               c("#2c83b7","deepskyblue1", "white",
                                 "red", "#c2395a"))

p <- Heatmap(plotdata, 
             col = col_fun,
             top_annotation = ht, 
             show_row_names = F,
             show_row_dend = F,
             show_column_dend = T,
             show_column_names = F,
             cluster_columns = T,
             clustering_method_columns = "ward.D",
             clustering_method_rows = "ward.D",
             cluster_rows = T,
             row_names_gp = gpar(fontsize = 6),
             heatmap_legend_param = list(title = "Zscore"))

```

#### NAT KW-DEP corrplot
```{r}
library(tibble)
library(Hmisc)
library(corrplot)
prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1, rowNames = F)

NAT_DEP_list <- openxlsx::read.xlsx("./2022proteome/NAT-KW-test_in_proteome.xlsx", sheet = 2, rowNames = F)

NAT_DEP_exp <- merge(NAT_DEP_list[1], prodata[,1:40], by="Name")
NAT_DEP_exp <- column_to_rownames(NAT_DEP_exp, var = "Name")

res <- rcorr(as.matrix(log2(1+NAT_DEP_exp)))

col1 <- colorRampPalette(c( "black","#00007f","blue","#007fff","cyan",
  "#7fff7f","#fff5a5","#ffaa64","#ff8264","#ff6464","#930077")) 

corrplot(res$r, order = "hclust",col=col1(300), col.lim=c(0.8,1),
         hclust.method = "ward.D",method = "circle",diag=FALSE,
         tl.col = "black", type = "upper", tl.pos = "lt")
corrplot(res$r, add = T,order = "hclust",col=col1(300), col.lim=c(0.8,1),
         hclust.method = "ward.D", number.digits = 2,number.cex = 0.3,
         method = "number",type = "lower", tl.col = "black",diag=FALSE,
         tl.pos = "n", cl.pos="n")   

corrplot(res$r, order = "hclust",col=col1(300), col.lim=c(0.8,1),
         hclust.method = "ward.D",method = "circle",
         tl.col = "black", tl.pos = "lt", addrect = 2)

library(factoextra)
library(igraph)

d <- dist(t(NAT_DEP_exp))

fit <- hclust(d, method = "ward.D2")
plot(fit, hang=-1, main="")

fviz_dend(fit, k=2, rect = T, rect_fill = T,
          rect_border = c("#00b8a9","#f6416c"))


```

### Corplot_NAT
```{r}
library(factoextra)
library(igraph)
prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1, rowNames = T)

NAT_exp <- as.data.frame(log2(prodata[,1:39]+1))
d <- dist(t(NAT_exp))

fit <- hclust(d, method = "ward.D2")
plot(fit, hang=-1, main="")

fviz_dend(fit, k=2, rect = T, rect_fill = T,
          rect_border = c("#00b8a9","#f6416c"))

```

####Corplot_QC for metabolome and lipidome
```{r}
library(Hmisc)
library(corrplot)
QC_data <- read.csv("./2022ProteomeMetabolism/QC for lipidome.csv", header = T, row.names = 1)

plotdata <- as.data.frame(log2(QC_data+1))

res <- rcorr(as.matrix(plotdata))

corrplot(res$r, order = "hclust", 
         hclust.method = "ward.D2",method = "circle",
         tl.col = "black", type = "upper", tl.pos = "d")

corrplot(res$r, add = T,order = "hclust", 
         hclust.method = "ward.D2",
         method = "number",type = "lower", tl.col = "black",diag=FALSE,
         tl.pos = "n", cl.pos="n")              


```

#### Hierarchical_Clustering_NAT_DEP
```{r}
library(factoextra)
library(igraph)

d <- dist(t(plotdata))

fit <- hclust(d, method = "complete")
plot(fit, hang=-1, main="")

fviz_dend(fit, k=2, rect = T, rect_fill = T,
          rect_border = c("#00b8a9","#f6416c"))
```

### NAT_LDA
```{r}
library(MASS)
library(ggplot2)
library(MOVICS)
prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1, rowNames = T)
pt.elite <- getElites(dat=prodata[,1:39],
                       method="mad",
                       na.action="rm",
                       elite.pct = 0.3)
pt.data <- as.data.frame(scale(t(log2(pt.elite$elite.dat+1))))

#pt.data <- as.data.frame(scale(t(log2(1+NAT_DEP_exp))))
lda.data <- as.data.frame(t(na.omit(t(pt.data))))

groups <- as.factor(c(rep("S1",3),rep("S2",5),rep("S3",5),
                          rep("S4",5),rep("S5",5),rep("S6",5),
                          rep("S7",5),rep("S8",6)))

lda.pt = lda(lda.data, groups)

result = predict(lda.pt, lda.data)

table(groups, result$class)

P = lda.pt$scaling 
means = lda.pt$means %*% P

total_means = as.vector(lda.pt$prior %*% means)
n_samples = nrow(lda.data)

x<-as.matrix(lda.data) %*% P - (rep(1, n_samples) %o% total_means)

subtybes <- as.data.frame(c(rep("Subtybe1",13),rep("Subtybe2",26)))
names(groups) <- "Groups"
names(subtybes) <- "Subtybes"
lda_data <- cbind(x, groups, subtybes)

p <- ggplot(data = lda_data, aes(x = LD1, y = LD2)) +
  geom_point(aes(color = groups, shape=groups), size = 2) +
  scale_shape_manual(values = c(0:2,5,6,3,4,8))+
  scale_color_manual(values = c("#48466d","#3d84a8","#46cdcf","#abedd8",
                                "#f9ed69","#f08a5d","#b83b5e","#6a2c70")) + 
  theme_bw() +
  theme(legend.position = "",
        panel.grid=element_blank(),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14)) +
  labs(x =  paste('LD1'), y = paste('LD2'), color = '')

g <- p + stat_ellipse(aes(fill = Subtybes), geom = 'polygon', level = 0.95, 
                 alpha = 0.1, show.legend = FALSE) +
    scale_fill_manual(values = c("#0f4392","#ff5151"))

```

### NAT_Lipds_OPLS
```{r}
library(MASS)
library(ggplot2)
library(ropls)
lipid.data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Lipid2023.xlsx", sheet = 1, rowNames = T)

nat.data <- as.data.frame(log2(lipid.data[,c(2:40)]+1))

pls.data <- as.data.frame(t(nat.data))

groups <- openxlsx::read.xlsx("./2022proteome/Clinicaldata2.xlsx", sheet = 4, rowNames = T)

##PLS-DA
pdf("NAT_Lipid.plsda.pdf")
pt.plsda <- opls(as.matrix(pls.data), groups$Subtypes2)
dev.off()

plot(pt.plsda,
     typeVc = "x-score",
     parAsColFcVn = groups$Subtypes2,
     parPaletteVc = c("#00b8a9","#f6416c"))

## OPLS-DA

pdf("NAT_Lipid.oplsda.pdf")
lipid.oplsda <- opls(as.matrix(pls.data), groups$Subtypes2, predI = 1, orthoI = NA)
dev.off()

plot(lipid.oplsda,
     typeVc = "x-score",
     parAsColFcVn = groups$Subtypes2,
     parPaletteVc = c("#00b8a9","#f6416c"))
```

### NAT_Metabolites_OPLS
```{r}
library(MASS)
library(ggplot2)
library(ropls)
lipid.data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolites2.xlsx", sheet = 2, rowNames = T)

pt.data <- as.data.frame(log2(lipid.data[,c(1:39)]+1))

pls.data <- as.data.frame(t(pt.data))

groups <- openxlsx::read.xlsx("./2022proteome/Clinicaldata2.xlsx", sheet = 4, rowNames = T)

## OPLS-DA

pdf("NAT_Metabolites.oplsda2.pdf")
lipid.oplsda <- opls(as.matrix(pls.data), groups$Subtypes2, predI = 1, orthoI = NA)
dev.off()

plot(lipid.oplsda,
     typeVc = "x-score",
     parAsColFcVn = groups$Subtypes2,
     parPaletteVc = c("#00b8a9","#f6416c"))
```

#### ggplot2 
```{r}
library(ggplot2)
library(ggsci)
library(tidyverse)

sample.score = lipid.oplsda@scoreMN %>%  
  as.data.frame() %>%
  mutate(group = groups$Subtypes2,
         o1 = lipid.oplsda@orthoScoreMN[,1]) 
head(sample.score)

p <- ggplot(sample.score, aes(p1, o1, color = groups$Segments)) +
  geom_hline(yintercept = 0, linetype = 'dashed', size = 0.5) + 
  geom_vline(xintercept = 0, linetype = 'dashed', size = 0.5) +
  geom_point(aes(color = groups$Segments, shape=groups$Segments), size = 2) +
  scale_shape_manual(values = c(0:2,5,6,3,4,8))+
  scale_color_manual(values = c("#48466d","#3d84a8","#46cdcf","#abedd8",
                                "#f9ed69","#f08a5d","#b83b5e","#6a2c70")) + 
  #geom_point(aes(-10,-10), color = 'white') +
  labs(x = 'OPLS-DA:t1(7%)',y = 'OPLS-DA:to1') +
  xlim(-30,30)+ylim(-80,60)+
  theme_bw() +
  theme(legend.position ="",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))

g <- ggplot(sample.score, aes(p1, o1, color = groups$Subtypes2)) +
  geom_hline(yintercept = 0, linetype = 'dashed', size = 0.5) + 
  geom_vline(xintercept = 0, linetype = 'dashed', size = 0.5) +
  geom_point() +
  labs(x = 'OPLS-DA:t1(7%)',y = 'OPLS-DA:to1') +
  stat_ellipse(level = 0.95, linetype = 'solid', 
               size = 1, show.legend = FALSE) +
  scale_color_manual(values = c("#0f4392","#ff5151")) +
  xlim(-30,30)+ylim(-80,60)+
  theme_bw() +
  theme(legend.position = c(0.11,0.93),
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))
```

##### I
```{r}
lipid.vip <- as.data.frame(getVipVn(lipid.oplsda))
write.csv(lipid.vip, file = "NAT_Metabolites_DAL_VIP.csv", row.names = T)
```


### DEP in NAT Clusters
```{r}
library(dplyr)
library(limma)
prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1,rowNames = T)

groups <- c(rep("Cluster1",13), rep("Cluster2",26))

design<- model.matrix(~0+factor(groups))
colnames(design)<- levels(factor(groups))
##voom
v <- voom(prodata[,1:39], design, plot=TRUE)

fit <- lmFit(v, design)

cont.matrix <- makeContrasts('Cluster1-Cluster2', levels = design) 
fit2 <- contrasts.fit(fit, cont.matrix)
fit2 <- eBayes(fit2)

output <- topTable(fit2, coef = 1, number = Inf, adjust.method = 'BH')
write.csv(output, file = "NAT_DEPs_limmavoom2.csv",row.names = T)

```

>Lipids

```{r}
library(dplyr)
library(limma)
lipid.data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Lipid2023.xlsx", sheet = 1,rowNames = T)%>% as.data.frame()
nat.data <- as.data.frame(lipid.data[,c(2:40)]+1)

groups <- c(rep("Cluster1",18), rep("Cluster2",21))

design<- model.matrix(~0+factor(groups))
colnames(design)<- levels(factor(groups))
##voom
v <- voom(nat.data, design, plot=TRUE)

fit <- lmFit(v, design)

cont.matrix <- makeContrasts('Cluster1-Cluster2', levels = design) 
fit2 <- contrasts.fit(fit, cont.matrix)
fit2 <- eBayes(fit2)

output <- topTable(fit2, coef = 1, number = Inf, adjust.method = 'BH')
write.csv(output, file = "NAT_DALs_limmavoom.csv",row.names = T)

```

>Metabolites

```{r}
metabolites.data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolites2.xlsx", sheet = 2,rowNames = T)

groups <- c(rep("Cluster1",18), rep("Cluster2",21))

design<- model.matrix(~0+factor(groups))
colnames(design)<- levels(factor(groups))
##voom
v <- voom(metabolites.data[,1:39], design, plot=TRUE)

fit <- lmFit(v, design)

cont.matrix <- makeContrasts('Cluster1-Cluster2', levels = design) 
fit2 <- contrasts.fit(fit, cont.matrix)
fit2 <- eBayes(fit2)

output <- topTable(fit2, coef = 1, number = Inf, adjust.method = 'BH')
write.csv(output, file = "NAT_DAMs_limmavoom3.csv",row.names = T)
```

### Heatmap_NAT_hCluster2
```{r}
library(tibble)
library(dplyr)
library(ComplexHeatmap)

PT_list <- openxlsx::read.xlsx("./NAT_DEPs_limmavoom2.xlsx", sheet = 2, rowNames = F)
PT_list <- as.data.frame(PT_list[,1])
names(PT_list) <- "Name" 

PT_list <- PT_list%>% distinct(Name, .keep_all = T)

prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1, rowNames = F)
## Read into KW test_expression matrix
dep.data <- merge(PT_list, prodata[,1:40], by = "Name")
dep.data <- column_to_rownames(dep.data, var = "Name")

plotdata <- as.data.frame(t(scale(log2(0.5+t(dep.data)))))

clinicaldata <- openxlsx::read.xlsx("./2022proteome/Clinicaldata2.xlsx", sheet = 4, rowNames = T)

ht = HeatmapAnnotation(Segments = clinicaldata$Segments,
                        show_legend = T,annotation_name_side = "left",
                        col = list(Segments = c("S1" = "#48466d",
                                               "S2"   = "#3d84a8",
                                               "S3"   = "#46cdcf",
                                               "S4"   = "#abedd8",
                                               "S5"   = "#f9ed69",
                                               "S6"   = "#f08a5d",
                                               "S7"   = "#b83b5e",
                                               "S8"   = "#6a2c70")))
ht2 = HeatmapAnnotation(Clusters = clinicaldata$Subtypes2,
                        show_legend = T,annotation_name_side = "left",
                        col = list(Clusters = c("Cluster1" = "#0f4392",
                                               "" = "#ff5151"))) 
col_fun = circlize::colorRamp2(c(-4,-2, 0,2, 4),
                               c("#49a09d","#00AFBB","white",
                                 "#FC4E07", "#b20a2c"))
set.seed(123)
p <- Heatmap(plotdata, column_split=clinicaldata$Subtypes2,
             col = col_fun,
             top_annotation = c(ht2,ht),
             show_row_names = F,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = F,clustering_distance_columns = "euclidean",
             cluster_columns = T, clustering_method_columns = "ward.D",
             cluster_rows = T, 
             row_names_gp = gpar(fontsize = 6),
             heatmap_legend_param = list(title = "Zscores"))


cls <- plyr::ldply (row_order(p), data.frame)
names(cls) <- c("id", "rowid") 
cls <- dplyr::arrange(cls, rowid)
plotdata$cluster <- cls$id
write.csv(plotdata, file = "Heatmap_PT_DEP_cluster.csv")

```

### Heatmap of DAMs and DALs in NAT
```{r}
library(tibble)
library(dplyr)
library(ComplexHeatmap)

dam_list <- openxlsx::read.xlsx("./NAT_DAMs_limmavoom2.xlsx", sheet = 2, rowNames = F)
dam_list <- as.data.frame(dam_list[,1])
names(dam_list) <- "Name" 

dam_list <- dam_list%>% distinct(Name, .keep_all = T)

metabolites.data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolites2.xlsx", sheet = 2, rowNames = F)
## Read into KW test_expression matrix
dep.data <- merge(dam_list, metabolites.data[,1:40], by = "Name")
dep.data <- column_to_rownames(dep.data, var = "Name")

plotdata <- as.data.frame(t(scale(log2(t(dep.data)))))

clinicaldata <- openxlsx::read.xlsx("./2022proteome/Clinicaldata2.xlsx", sheet = 4, rowNames = T)

ht = HeatmapAnnotation(Segments = clinicaldata$Segments,
                        show_legend = T,annotation_name_side = "left",
                        col = list(Segments = c("S1" = "#48466d",
                                               "S2"   = "#3d84a8",
                                               "S3"   = "#46cdcf",
                                               "S4"   = "#abedd8",
                                               "S5"   = "#f9ed69",
                                               "S6"   = "#f08a5d",
                                               "S7"   = "#b83b5e",
                                               "S8"   = "#6a2c70")))
ht2 = HeatmapAnnotation(Clusters = clinicaldata$Subtypes2,
                        show_legend = T,annotation_name_side = "left",
                        col = list(Clusters = c("Cluster1" = "#0f4392",
                                               "Cluster2" = "#ff5151"))) 
col_fun = circlize::colorRamp2(c(-4,-2, 0,2, 4),
                               c("#49a09d","#00AFBB","white",
                                 "#FC4E07", "#b20a2c"))
set.seed(123)
p <- Heatmap(plotdata, 
             col = col_fun,
             top_annotation = c(ht2,ht),
             show_row_names = F,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = F,
             cluster_columns = F, 
             cluster_rows = T, 
             row_names_gp = gpar(fontsize = 6),
             heatmap_legend_param = list(title = "Zscores"))
```

> lipids

```{r}
library(tibble)
library(dplyr)
library(ComplexHeatmap)

dal_list <- openxlsx::read.xlsx("./LL-RL-NAT/NAT_DALs_limmavoom.xlsx", sheet = 2, rowNames = F)
dal_list <- as.data.frame(dal_list[,1])
names(dal_list) <- "Name" 

dal_list <- dal_list%>% distinct(Name, .keep_all = T)

lipids.data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Lipid2023.xlsx", sheet = 1, rowNames = F)
## Read into KW test_expression matrix
dep.data <- merge(dal_list, lipids.data[,c(1:41)], by = "Name")
dep.data <- column_to_rownames(dep.data[,-2], var = "Name")

plotdata <- as.data.frame(t(scale(log2(t(dep.data)))))

clinicaldata <- openxlsx::read.xlsx("./2022proteome/Clinicaldata2.xlsx", sheet = 4, rowNames = T)

ht = HeatmapAnnotation(Segments = clinicaldata$Segments,
                        show_legend = T,annotation_name_side = "left",
                        col = list(Segments = c("S1" = "#48466d",
                                               "S2"   = "#3d84a8",
                                               "S3"   = "#46cdcf",
                                               "S4"   = "#abedd8",
                                               "S5"   = "#f9ed69",
                                               "S6"   = "#f08a5d",
                                               "S7"   = "#b83b5e",
                                               "S8"   = "#6a2c70")))
ht2 = HeatmapAnnotation(Clusters = clinicaldata$Subtypes2,
                        show_legend = T,annotation_name_side = "left",
                        col = list(Clusters = c("Cluster1" = "#0f4392",
                                               "Cluster2" = "#ff5151"))) 
col_fun = circlize::colorRamp2(c(-4,-2, 0,2, 4),
                               c("#49a09d","#00AFBB","white",
                                 "#FC4E07", "#b20a2c"))
set.seed(123)
p <- Heatmap(plotdata, 
             col = col_fun,
             top_annotation = c(ht2,ht),
             show_row_names = F,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = F,clustering_distance_columns = "euclidean",
             cluster_columns = F, clustering_method_columns = "ward.D",
             cluster_rows = T, 
             row_names_gp = gpar(fontsize = 6),
             heatmap_legend_param = list(title = "Zscores"))
```

#### Bubble plot of MSEA in NAT-cluster
```{r}
library(ggplot2)

ll.data <- read.csv("./2022ProteomeMetabolism/NAT-Cluster1-msea_ora_result.csv", header = T)

plotdata <- ll.data[c(1:8),c(1,8,9)]

p <- ggplot(plotdata,aes(x = EnrichmentRatio, 
              y = reorder(Name,EnrichmentRatio,sum), 
              size = EnrichmentRatio,
              colour=LogP)) +
     geom_point(shape = 16) +                
     labs(x = "EnrichmentRatio", y = "")+          
     scale_colour_gradient2(name="-log10(P-value)",
     low = "#ffbd39",
     mid = "#e61c5d",
     high = "#930077",
     midpoint = 3)+
     theme_bw()          


RL.data <- read.csv("./2022ProteomeMetabolism/NAT-cluster2-msea_ora_result.csv", header = T)

plotdata <- RL.data[c(1:8),c(1,8,9)]

```

#### LineChart_NAT_DAL
```{r}
library(ggplot2)
library(reshape2)
library(tibble)
library(dplyr)
hex.data <- openxlsx::read.xlsx("./NAT/NAT_lipid_mean2023.xlsx", sheet = 9, rowNames = T)

hex.data <- as.data.frame(t(scale(t(hex.data))))
hex.data <- rownames_to_column(hex.data, var = "Name")

plotdata <- melt(hex.data, by="Name")

mean <- plotdata %>% 
  group_by(variable) %>% 
  summarise(value = mean(value))  
mean$variable <- as.numeric(mean$variable)

p <- ggplot(plotdata, aes(x=variable, y=value)) +
     geom_line(aes(group=Name), color="#e84545", alpha=0.03, size=2)+
     geom_line(data = mean, color="#e84545", alpha = 0.7, size=2) + 
     geom_point(data = mean)+
     xlab("")+ylab("Relative Abundance")+
     theme_bw()

```


### NAT_annotation
```{r}
library(ggplot2)
cluster1 <- openxlsx::read.xlsx("./NAT_hcluster_FINAL_GO.xlsx", sheet = 2)

c1 <- ggplot(cluster1, aes(x = reorder(factor(Description), `-LogP`), 
                           y = `-LogP`)) + 
  geom_bar(stat = 'identity', fill = '#0f4392', width = 0.7)+
  xlab("")+ylab("-log10(p-value)")+
  coord_flip()+
  theme_classic() +
  theme(legend.position ="",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))

cluster2 <- openxlsx::read.xlsx("./NAT_hcluster_FINAL_GO.xlsx", sheet = 3)
c2 <- ggplot(cluster2, aes(x = reorder(factor(Description), `-LogP`), 
                           y = `-LogP`)) + 
  geom_bar(stat = 'identity', fill = '#ff5151', width = 0.7)+
  xlab("")+ylab("-log10(p-value)")+
  coord_flip()+
  theme_classic() +
  theme(legend.position ="",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))


```

##Multi-Omics
#### T/N data LDA
```{r}
library(MASS)
library(ggplot2)
library(MOVICS)
prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 2, rowNames = T)
pt.elite <- getElites(dat=prodata,
                       method="mad",
                       na.action="rm",
                       elite.pct = 0.3)
pt.data <- as.data.frame(scale(t(log2(pt.elite$elite.dat+1))))

lda.data <- as.data.frame(t(na.omit(t(pt.data))))

groups <- as.factor(c(rep("S1",3),rep("S2",5),rep("S3",5),
                          rep("S4",5),rep("S5",5),rep("S6",5),
                          rep("S7",5),rep("S8",6)))

lda.pt = lda(lda.data, groups)

result = predict(lda.pt, lda.data)

table(groups, result$class)

P = lda.pt$scaling 
means = lda.pt$means %*% P

total_means = as.vector(lda.pt$prior %*% means)
n_samples = nrow(lda.data)

x<-as.matrix(lda.data) %*% P - (rep(1, n_samples) %o% total_means)

subtybes <- as.data.frame(c(rep("Subtybe1",18),rep("Subtybe2",21)))
names(groups) <- "Groups"
names(subtybes) <- "Subtybes"
lda_data <- cbind(x, groups, subtybes)

p <- ggplot(data = lda_data, aes(x = LD1, y = LD2)) +
  geom_point(aes(color = groups, shape=groups), size = 2) +
  scale_shape_manual(values = c(0:2,5,6,3,4,8))+
  scale_color_manual(values = c("#48466d","#3d84a8","#46cdcf","#abedd8",
                                "#f9ed69","#f08a5d","#b83b5e","#6a2c70")) + 
  theme_bw() +
  theme(legend.position = "",
        panel.grid=element_blank(),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14)) +
  labs(x =  paste('LD1'), y = paste('LD2'), color = '')

g <- p + stat_ellipse(aes(fill = Subtybes), geom = 'polygon', level = 0.95, 
                 alpha = 0.1, show.legend = FALSE) +
    scale_fill_manual(values = c("#00b8a9","#f6416c"))

```


### proteome NMF cluster

```{r}
library(NMF)
library(doMPI)
library(dplyr)
prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 2, rowNames = T)%>%as.data.frame()

prodata <- log2(1+prodata)

mads <- apply(prodata,1,mad) 
pro.exp <- prodata[rev(order(mads))[1:4000],] 

res <- nmf(pro.exp,2:8,nrun=10,seed=123,method = "lee")
plot(res)

res <- nmf(pro.exp,3,nrun=10,seed=123)

group <- predict(res)
table(group)
write.csv(as.data.frame(group), file = "Proteome_subtypes4000.csv")

coefmap(res)
consensusmap(res)

index <- extractFeatures(res,0.95) 
sig.order <- unlist(index)
NMF.Exp <- prodata[sig.order,]
NMF.Exp <- na.omit(NMF.Exp)

library(tinyarray)
dp = NMF.Exp[,order(group)]
draw_heatmap(dp,sort(group), split_column = 3,
             color_an = c("#2874C5","#EABF00","#C6524A","#868686"),
             annotation_legend = T,
             cluster_cols = F,
             show_rownames = T)


jco <- c("#2874C5","#EABF00","#C6524A")
res2 <- nmf(NMF.Exp,3,nrun=10,seed=123)

coefmap(res2)
consensusmap(res2)
group2 <- predict(res2)

basismap(res, distfun="euclidean", hclustfun="ward",
 cexCol = 1,
 cexRow = 0.3,
 annColors=list(c("1"=jco[1],"2"=jco[2],"3"=jco[3])))

group <- predict(res) 
table(group)
consensusmap(res,
             labRow = NA,
             labCol = NA,
             annCol = data.frame("cluster"=group[colnames(res)]),
             annColors = list(cluster=c("1"=jco[1],"2"=jco[2],"3"=jco[3])))

```

#### lipid NMF cluster

```{r}
library(NMF)
library(doMPI)
library(dplyr)
lipid.data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolites2023.xlsx", sheet = 3, rowNames = T)%>%as.data.frame()

lipid.data <- log2(1+lipid.data)
mads <- apply(lipid.data,1,mad) 
lipid.exp <- lipid.data[rev(order(mads))[1:500],] 

res <- nmf(lipid.exp,2:8,nrun=10,seed=123,method = "lee")
plot(res)

res <- nmf(lipid.data,3,nrun=10,seed=123)

group <- predict(res)
table(group)
write.csv(as.data.frame(group), file = "Proteome_subtypes4000.csv")

coefmap(res)
consensusmap(res)

index <- extractFeatures(res,0.95) 
sig.order <- unlist(index)
NMF.Exp <- prodata[sig.order,]
NMF.Exp <- na.omit(NMF.Exp) 

library(tinyarray)
dp = NMF.Exp[,order(group)]
draw_heatmap(dp,sort(group), split_column = 3,
             color_an = c("#2874C5","#EABF00","#C6524A","#868686"),
             annotation_legend = T,
             cluster_cols = F,
             show_rownames = T)


jco <- c("#2874C5","#EABF00","#C6524A")
res2 <- nmf(NMF.Exp,3,nrun=10,seed=123)

coefmap(res2)
consensusmap(res2)
group2 <- predict(res2)

basismap(res, distfun="euclidean", hclustfun="ward",
 cexCol = 1,
 cexRow = 0.3,
 annColors=list(c("1"=jco[1],"2"=jco[2],"3"=jco[3])))

group <- predict(res) 
table(group)
consensusmap(res,
             labRow = NA,
             labCol = NA,
             annCol = data.frame("cluster"=group[colnames(res)]),
             annColors = list(cluster=c("1"=jco[1],"2"=jco[2],"3"=jco[3])))

```


#### proteome 3 clusters
```{r}
library(limma)
prodata <- read.csv("./NMF_log2_mad3000.csv", header = T, row.names = 1)
groups <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 3, rowNames = TRUE)
groups <- groups[1:39,]
design<- model.matrix(~0+factor(groups$Subtypes3))
colnames(design)<- levels(factor(groups$Subtypes3))

fit <- lmFit(prodata, design)

cont.matrix <- makeContrasts('C3-other', levels = design) 
fit2 <- contrasts.fit(fit, cont.matrix)
fit2 <- eBayes(fit2)

output <- topTable(fit2, coef = 1, number = Inf, adjust.method = 'BH')
write.csv(output, file = "NMF_C3_DEPs_mad.csv",row.names = T)
```

>Multiple analysis

```{r}
library(limma)
library(dplyr)
##Proteome
prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 2, rowNames = T)%>%as.data.frame()
prodata <- as.data.frame(log2(1+prodata))

groups <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 3, rowNames = TRUE)
groups <- groups[1:39,]
design<- model.matrix(~0+factor(groups$Subtypes))
colnames(design)<- levels(factor(groups$Subtypes))

cont.wt <- makeContrasts("C1-C2", "C1-C3","C2-C3", "C2-C1",
                         levels=design) 

fit <- lmFit(prodata, design)
fit2 <- contrasts.fit(fit, cont.wt) 
fit2 <- eBayes(fit2) 
tT=topTableF(fit2, adjust="BH",sort.by="F",n=Inf)

write.csv(tT, file = "NMF_C1-2-3_DEPs.csv",row.names = T)

##metabolome 
metabo.data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolites2023.xlsx", sheet = 5, rowNames = T)%>%as.data.frame()
metabo.data <- as.data.frame(log2(1+metabo.data))

groups <- openxlsx::read.xlsx("./2022proteome/Clinicaldata2.xlsx", sheet = 3, rowNames = TRUE)
groups <- groups[1:39,]
design<- model.matrix(~0+factor(groups$Subtypes))
colnames(design)<- levels(factor(groups$Subtypes))

cont.wt <- makeContrasts("C1-C2", "C1-C3","C2-C3", "C2-C1",
                         levels=design) 

fit <- lmFit(metabo.data, design)
fit2 <- contrasts.fit(fit, cont.wt) 
fit2 <- eBayes(fit2) 
tT=topTableF(fit2, adjust="BH",sort.by="F",n=Inf)

write.csv(tT, file = "NMF_C1-2-3_DAMs.csv",row.names = T)

## lipidome
lipid.data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Lipid2023.xlsx", sheet = 3, rowNames = T)%>%as.data.frame()
lipid.data <- as.data.frame(log2(1+lipid.data))

groups <- openxlsx::read.xlsx("./2022proteome/Clinicaldata2.xlsx", sheet = 3, rowNames = TRUE)
design<- model.matrix(~0+factor(groups$Subtypes))
colnames(design)<- levels(factor(groups$Subtypes))

cont.wt <- makeContrasts("C1-C2", "C1-C3","C2-C3", "C2-C1",
                         levels=design) 

fit <- lmFit(lipid.data, design)
fit2 <- contrasts.fit(fit, cont.wt) 
fit2 <- eBayes(fit2) 
tT=topTableF(fit2, adjust="BH",sort.by="F",n=Inf)

write.csv(tT, file = "NMF_C1-2-3_DALs.csv",row.names = T)

```

#### NMF clusters anotations
```{r}
library(pheatmap)
plotdata <- openxlsx::read.xlsx("./NMF3_FINAL_GO.xlsx", sheet=3, rowNames = T)
p <- pheatmap(plotdata, cluster_cols = FALSE, cluster_rows = FALSE,
              border_color = "black",
              color = colorRampPalette(c("white","red","firebrick3"))(20))
```

#### NMF clusters heatmap
```{r}
library(tibble)
library(dplyr)
library(ComplexHeatmap)
library(circlize)
nmf_list <- openxlsx::read.xlsx("./NMF_C1-2-3_DEPs.xlsx", sheet = 2, rowNames = F)
nmf_list <- as.data.frame(nmf_list[,1])
names(nmf_list) <- "Name" 

nmf_list <- nmf_list%>% distinct(Name, .keep_all = T)

prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 2, rowNames = F)%>%as.data.frame()

dep.data <- merge(nmf_list, prodata, by = "Name", sort = FALSE)
dep.data <- column_to_rownames(dep.data, var = "Name")

plotdata <- as.data.frame(t(scale(log2(1+t(dep.data)))))

clinicaldata <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 2, rowNames = T)%>%as.data.frame()

sub = HeatmapAnnotation(Subtypes = clinicaldata$Subtypes,
                        show_legend = T,annotation_name_side = "left",
            col = list(Subtypes = c("C1" = "#2874C5",
                                    "C2" = "#EABF00",
                                    "C3" = "#C6524A")))
seg = HeatmapAnnotation(Segments = clinicaldata$Segments,
                        show_legend = T,annotation_name_side = "left",
             col = list(Segments  = c("S1" = "#48466d", "S2"   = "#3d84a8",
                                     "S3" = "#46cdcf", "S4"   = "#abedd8",
                                     "S5" = "#f9ed69", "S6"   = "#f08a5d",
                                     "S7" = "#b83b5e", "S8"   = "#6a2c70")))

col_lianxu1 <- colorRamp2(c(0, 3),  c("#b1cbfa", "#6643b5"))
col_lianxu2 <- colorRamp2(c(0, 4),  c("#b1cbfa", "#6643b5"))
col_lianxu3 <- colorRamp2(c(0, 40),  c("#b1cbfa", "#6643b5"))
col_lianxu4 <- colorRamp2(c(15, 35),  c("#b1cbfa", "#6643b5"))
ht = HeatmapAnnotation(Age      = clinicaldata$Age,
                            Gender   = clinicaldata$Gender,
                            Drinking = clinicaldata$Drinking,
                            HBV      = clinicaldata$HBV,
                            AFP      = clinicaldata$AFP_cut,
                            Grade    = clinicaldata$Grade,
                            TumorSize= clinicaldata$TumorDiameter,
                            BMI      = clinicaldata$BMI,
                            TBIL     = clinicaldata$TBIL,
                            `ALB/GLB`= clinicaldata$`ALB/GLB`,
                            `AST/ALT`= clinicaldata$`AST/ALT`,
                            show_legend = T,annotation_name_side = "left",
       col = list(Age = circlize::colorRamp2(breaks = c(min(clinicaldata$Age),                                                        median(clinicaldata$Age),
                                                        max(clinicaldata$Age)), 
                  colors = c("#b1cbfa", "#b1cbfa", "#6643b5")),
                  Gender = c("M" = "#6643b5", "F" = "#b1cbfa"),
                  Drinking=c("Yes"="#6643b5", "No"= "#b1cbfa"),
                  HBV  = c("Positive"="#6643b5", "Negative"= "#b1cbfa"),
                  AFP = c("10"="#dfe2fe","10-100"= "#b1cbfa",
                              "100-1000"="#8e98f5",">1000"="#6643b5"),
                  Grade = c("G1"="#dfe2fe","G1-G2"= "#b1cbfa", "G2"="#8e98f5",
                            "G2-G3"="#7874f2", "G3"="#6643b5"),
                  TumorSize=c(">5" = "#6643b5", "5" = "#b1cbfa"),
                  `ALB/GLB`=col_lianxu1, `AST/ALT`=col_lianxu2,
                  BMI=col_lianxu4, TBIL=col_lianxu3))

col_fun = circlize::colorRamp2(c(-4,-2, 0, 2, 4),
                               c("#49a09d","#00AFBB","white",
                                 "#FC4E07", "#b20a2c"))
set.seed(123)
p <- Heatmap(plotdata, km=3,
             col = col_fun, column_split = clinicaldata$Subtypes,
             top_annotation = c(sub, seg ,ht),
             show_row_names = F,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = F,
             cluster_columns = F, 
             cluster_rows = F, 
             row_names_gp = gpar(fontsize = 6),
             heatmap_legend_param = list(title = "Zscores"))
```

#### Ribosome biogenesis
```{r}
library(dplyr)
library(tibble)
prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1, rowNames = F)%>%as.data.frame()
rb.list <- openxlsx::read.xlsx("./NMF_Ribosome b_genes.xlsx", sheet = 6, rowNames = F, startRow = 5)
rb.list <- as.data.frame(rb.list[,1])
names(rb.list) <- "Name" 

rb.list <- rb.list%>% distinct(Name, .keep_all = T)

rb.data <- merge(rb.list, prodata, by="Name",sort = F)
write.csv(rb.data, file = "Ribosome b_genes.csv", row.names = F)

rb.data <- column_to_rownames(rb.data, var = "Name")
rb.data <- as.data.frame(t(rb.data))

clinicaldata <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 3, rowNames = T)%>%as.data.frame()

rb.aov <- cbind(clinicaldata$Subtypes, rb.data)
names(rb.aov)[1] <- "Subtypes"
rb.aov <- rb.aov[-c(1:39),]

anova1 <- summary(aov(rb.aov[,2]~Subtypes, rb.aov))

output <- as.data.frame(anova1[[1]])


for (i in 2:ncol(rb.aov)) {
  anova = summary(aov(rb.aov[,i]~Subtypes, rb.aov))
  output1 <- as.data.frame(anova[[1]])
  output <- rbind(output1,output)
}

write.csv(output, file="Annova.results.csv")

```

>heatmap

```{r}
library(dplyr)
library(ComplexHeatmap)
rb.mean <- openxlsx::read.xlsx("./NMF_Ribosome b_genes.xlsx", sheet = 7, rowNames = T)

plotdata <- as.data.frame(scale(t(rb.mean)))

col_fun = circlize::colorRamp2(c(-1.5,-0.5, 0, 0.5, 1.5),
                               c("#49a09d","#00AFBB","white",
                                 "#FC4E07", "#b20a2c"))

cell_fun <- function(j, i, x, y, width, height, fill) {
  grid.text(
    round(plotdata[i, j], 1), 
    x, y,
    gp = gpar(
      fontsize = 10
    ))
}

set.seed(123)
p <- Heatmap(plotdata,
             rect_gp = gpar(col = "white", lwd = 2), 
             col = col_fun, cell_fun = cell_fun,
             show_row_names = T,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = T,
             cluster_columns = F, 
             cluster_rows = F, 
             row_names_gp = gpar(fontsize = 10),
             heatmap_legend_param = list(title = "Zscores"))
```

#### Ribosome proteins
```{r}
library(dplyr)
library(ComplexHeatmap)
RP.data <- openxlsx::read.xlsx("./NMF_Ribosome b_genes.xlsx", sheet = 3, rowNames = T)%>%as.data.frame()

plotdata <- as.data.frame(t(scale(log2(1+t(RP.data)))))

clinicaldata <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 3, rowNames = T)%>%as.data.frame()

sub = HeatmapAnnotation(Subtypes = clinicaldata$Subtypes,
                        show_legend = T,annotation_name_side = "left",
            col = list(Subtypes = c("NAT"= "grey",
                                    "C1" = "#2874C5",
                                    "C2" = "#EABF00",
                                    "C3" = "#C6524A")))
seg = HeatmapAnnotation(Segments = clinicaldata$Segments,
                        show_legend = T,annotation_name_side = "left",
             col = list(Segments  = c("S1" = "#48466d", "S2"   = "#3d84a8",
                                     "S3" = "#46cdcf", "S4"   = "#abedd8",
                                     "S5" = "#f9ed69", "S6"   = "#f08a5d",
                                     "S7" = "#b83b5e", "S8"   = "#6a2c70")))
hlob = HeatmapAnnotation(Hepatic_lobes = clinicaldata$Hepatic_lobes,
                        show_legend = T,annotation_name_side = "left",
             col = list(Hepatic_lobes=c("Left"= "#0f4392","Right"= "#ff5151")))

group = HeatmapAnnotation(Groups = clinicaldata$Groups,
                        show_legend = T,annotation_name_side = "left",
             col = list(Groups=c("NAT"= "#0f4392","TT"= "#ff5151")))

mark_gene <- openxlsx::read.xlsx("./NMF_Ribosome b_genes.xlsx", sheet = 4)%>%
as.data.frame()

row_anno <-  rowAnnotation(link = anno_mark(at = which(rownames(plotdata) 
                                                       %in% mark_gene$Name), 
                                       labels = mark_gene$Name,
                           labels_gp = gpar(fontsize = 7.5)))


col_fun = circlize::colorRamp2(c(-4,-2, 0, 2, 4),
                               c("#49a09d","#00AFBB","white",
                                 "#FC4E07", "#b20a2c"))
set.seed(111)
p <- Heatmap(plotdata,
             col = col_fun, column_split = clinicaldata$Subtypes,
             top_annotation = c(sub,seg,hlob),
             show_row_names = F,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = F,
             cluster_columns = T, 
             cluster_rows = F, 
             row_names_gp = gpar(fontsize = 7.5),
             heatmap_legend_param = list(title = "Zscores"))
p+row_anno
```

#### Ribosome biogenesis and ribosome proteins
```{r}
library(ggplot2)
library(ggrepel)
output <- openxlsx::read.xlsx("./Proteome_DEGs_limmavoom.xlsx", sheet = 3, rowNames = T)
names(output) <- c('Groups','log2FC', 'AveExpr', 'Pvalue', 'BH-Pvalue')

yanse <- ifelse(output$`BH-Pvalue`<0.01 & abs(output$log2FC)>= 1, ifelse(output$log2FC> 1,'red','grey'),'green')

p <- ggplot(output, aes(output$log2FC, -log10(output$`BH-Pvalue`), col=Groups))+geom_point(size=1)

g <- p+theme_bw()+ 
  scale_color_manual(values = c('#a82ffc','#ef7e56'))+ 
  labs(x="log2 (Fold Change)",y="-log10 (BH-Pvalue)") +
  geom_hline(yintercept = -log10(0.05), lty=4,col="grey",lwd=0.6) +
  geom_vline(xintercept = c(-1, 1), lty=4,col="grey",lwd=0.6) +
  theme(legend.position = "none",
        panel.grid=element_blank(),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))
##label
label <- ifelse(-log10(output$`BH-Pvalue`) > 5 & abs(output$log2FC) >= 1, rownames(output),"")

l <- g+geom_text_repel(data=output, aes(x = output$log2FC, 
                                   y = -log10(output$`BH-Pvalue`), 
                                   label = label),
                  size = 3,box.padding = unit(0.2, "lines"),
                  segment.color = "black", 
                  show.legend = FALSE)
```

#### xCell for immune cells
```{r}
library(xCell)
library(dplyr)
library(ComplexHeatmap)
exprMatrix <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1, rowNames = T)%>%as.data.frame()

xcell.data <- as.data.frame(xCellAnalysis(exprMatrix))

plotdata <- as.data.frame(t(scale(log2(1+t(xcell.data)))))

plotdata <- na.omit(plotdata)

clinicaldata <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 3, rowNames = T)%>%as.data.frame()

sub = HeatmapAnnotation(Subtypes = clinicaldata$Subtypes,
                        show_legend = T,annotation_name_side = "left",
            col = list(Subtypes = c("NAT"= "grey",
                                    "C1" = "#2874C5",
                                    "C2" = "#EABF00",
                                    "C3" = "#C6524A")))
seg = HeatmapAnnotation(Segments = clinicaldata$Segments,
                        show_legend = T,annotation_name_side = "left",
             col = list(Segments  = c("S1" = "#48466d", "S2"   = "#3d84a8",
                                     "S3" = "#46cdcf", "S4"   = "#abedd8",
                                     "S5" = "#f9ed69", "S6"   = "#f08a5d",
                                     "S7" = "#b83b5e", "S8"   = "#6a2c70")))
hlob = HeatmapAnnotation(Hepatic_lobes = clinicaldata$Hepatic_lobes,
                        show_legend = T,annotation_name_side = "left",
             col = list(Hepatic_lobes=c("Left"= "#0f4392","Right"= "#ff5151")))

group = HeatmapAnnotation(Groups = clinicaldata$Groups,
                        show_legend = T,annotation_name_side = "left",
             col = list(Groups=c("NAT"= "#0f4392","TT"= "#ff5151")))

col_fun = circlize::colorRamp2(c(-4,-2, 0, 2, 4),
                               c("#49a09d","#00AFBB","white",
                                 "#FC4E07", "#b20a2c"))
set.seed(111)
p <- Heatmap(plotdata,
             col = col_fun, column_split = clinicaldata$Subtypes,
             top_annotation = c(sub,seg,hlob),
             show_row_names = T,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = F,
             cluster_columns = T, 
             cluster_rows = T, 
             row_names_gp = gpar(fontsize = 7.5),
             heatmap_legend_param = list(title = "Zscores"))
```

#### NMF-C3 lipid groups
```{r}
library(dplyr)
library(ComplexHeatmap)
lipid.group <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Lipid_headgroup.xlsx", sheet = 5, rowNames = T)%>%as.data.frame()

plotdata <- as.data.frame(t(scale(log2(1+lipid.group[,-1]))))

clinicaldata <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 3, rowNames = T)%>%as.data.frame()

sub = HeatmapAnnotation(Subtypes = clinicaldata$Subtypes,
                        show_legend = T,annotation_name_side = "left",
            col = list(Subtypes = c("NAT"= "grey",
                                    "C1" = "#2874C5",
                                    "C2" = "#EABF00",
                                    "C3" = "#C6524A")))

col_fun = circlize::colorRamp2(c(-4,-2, 0, 2, 4),
                               c("#49a09d","#00AFBB","white",
                                 "#FC4E07", "#b20a2c"))
set.seed(111)
p <- Heatmap(plotdata,
             col = col_fun, column_split = clinicaldata$Subtypes,
             top_annotation = sub,
             show_row_names = T,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = F,
             cluster_columns = F, 
             cluster_rows = T, 
             row_names_gp = gpar(fontsize = 7.5),
             heatmap_legend_param = list(title = "Zscores"))
```


#### Neutrophil degranulation
```{r}
library(dplyr)
prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 2, rowNames = TRUE)%>%as.data.frame()

clinicaldata <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 2, rowNames = T)%>%as.data.frame()

plotdata <- cbind(as.data.frame(t(prodata)),clinicaldata$Subtypes)

Csubtype.mean <- group_by(plotdata, `clinicaldata$Subtypes`)%>%
  summarise_each(funs(median))

mean.data <- as.data.frame(t(Csubtype.mean))

openxlsx::write.xlsx(mean.data, file = "./NMF_Csubtypes_median.xlsx", rowNames=T)
```

>heatmap

```{r}
library(ComplexHeatmap)
ne.data <- openxlsx::read.xlsx("./NMF_C1-2-3_DEPs.xlsx", sheet = 3, rowNames = T)
plotdata <- as.data.frame(ne.data[,12:14])
  

col_fun = circlize::colorRamp2(c(-10, -5, -2, 0, 5, 10, 15, 20),
                               c("#6643b5", "#1565c0","#009fff", "#fffde4",
                                 "#fbd786", "#f7797d",
                                 "#ec2f4b", "#8a2387"))

cell_fun <- function(j, i, x, y, width, height, fill) {
  grid.text(
    round(plotdata[i, j], 1), 
    x, y,
    gp = gpar(
      fontsize = 10
    ))
}

set.seed(123)
p <- Heatmap(plotdata,  
             rect_gp = gpar(col = "white", lwd = 2, width=12), 
             col = col_fun, cell_fun = cell_fun,
             show_row_names = T,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = T,
             cluster_columns = F, 
             cluster_rows = F, 
             row_names_gp = gpar(fontsize = 6),
             heatmap_legend_param = list(title = "log2(T/N)"))
```

>heatmap2

```{r}
library(dplyr)
library(tibble)
library(circlize)
library(ComplexHeatmap)
ne.list <- openxlsx::read.xlsx("./2022proteome/Neutrophil granule proteins.xlsx", sheet = 2, rowNames = F)%>%as.data.frame()

ne.list <- as.data.frame(ne.list[,1])
names(ne.list) <- "Name" 

ne.list <- ne.list%>% distinct(Name, .keep_all = T)

prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1, rowNames = F)%>%as.data.frame()

ne.data <- merge(ne.list, prodata, by="Name",sort = F)
ne.data <- column_to_rownames(ne.data, var = "Name")

plotdata <- as.data.frame(t(scale(log2(1+t(ne.data)))))

clinicaldata <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 3, rowNames = T)%>%as.data.frame()

col_lianxu1 <- colorRamp2(c(0, 0.05,0.1),  c("#f2fcfe","#6dd5fa","#2980b9"))
col_lianxu2 <- colorRamp2(c(0, 0.1,0.2),  c("#fffcdc","#ffaf7b", "#d76d77"))

ht = HeatmapAnnotation(Neutrophils    = clinicaldata$Neutrophils,
                       ImmuneScores   = clinicaldata$ImmuneScore,
                         show_legend  = T, annotation_name_side = "left",
       col = list(Neutrophils=col_lianxu1, ImmuneScores=col_lianxu2))

sub = HeatmapAnnotation(Subtypes = clinicaldata$Subtypes,
                        show_legend = T,annotation_name_side = "left",
            col = list(Subtypes = c("NAT"= "grey",
                                    "C1" = "#2874C5",
                                    "C2" = "#EABF00",
                                    "C3" = "#C6524A")))
seg = HeatmapAnnotation(Segments = clinicaldata$Segments,
                        show_legend = T,annotation_name_side = "left",
             col = list(Segments  = c("S1" = "#48466d", "S2"   = "#3d84a8",
                                     "S3" = "#46cdcf", "S4"   = "#abedd8",
                                     "S5" = "#f9ed69", "S6"   = "#f08a5d",
                                     "S7" = "#b83b5e", "S8"   = "#6a2c70")))
hlob = HeatmapAnnotation(Hepatic_lobes = clinicaldata$Hepatic_lobes,
                        show_legend = T,annotation_name_side = "left",
             col = list(Hepatic_lobes=c("Left"= "#0f4392","Right"= "#ff5151")))

col_fun = circlize::colorRamp2(c(-4,-2, 0, 2, 4),
                               c("#49a09d","#00AFBB","white",
                                 "#FC4E07", "#b20a2c"))
set.seed(111)
p <- Heatmap(plotdata,
             column_dend_reorder = order(clinicaldata$ImmuneScore, decreasing = TRUE),
             col = col_fun,column_split = clinicaldata$Subtypes,
             top_annotation = c(ht,sub,seg,hlob),
             show_row_names = T,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = F,
             cluster_columns = T, 
             cluster_rows = T, 
             row_names_gp = gpar(fontsize = 7.5),
             heatmap_legend_param = list(title = "Zscores"))

```

##### NMF clusters heatmap in Metabolome and lipidome
```{r}
library(tibble)
library(dplyr)
library(ComplexHeatmap)
library(circlize)

nmf_list <- openxlsx::read.xlsx("./2022ProteomeMetabolism/NMF_C1-2-3_DALs.xlsx", sheet = 2, rowNames = F)
nmf_list <- as.data.frame(nmf_list[,1])
names(nmf_list) <- "Name" 

nmf_list <- nmf_list%>% distinct(Name, .keep_all = T)

prodata <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Lipid2023.xlsx", sheet = 3, rowNames = F)%>%as.data.frame()

dep.data <- merge(nmf_list, prodata, by = "Name", sort = FALSE)
dep.data <- column_to_rownames(dep.data, var = "Name")

plotdata <- as.data.frame(t(scale(log2(1+t(dep.data)))))

clinicaldata <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 2, rowNames = T)%>%as.data.frame()

sub = HeatmapAnnotation(Subtypes = clinicaldata$Subtypes,
                        show_legend = T,annotation_name_side = "left",
            col = list(Subtypes = c("C1" = "#2874C5",
                                    "C2" = "#EABF00",
                                    "C3" = "#C6524A")))

col_fun = circlize::colorRamp2(c(-4,-2, 0, 2, 4),
                               c("#49a09d","#00AFBB","white",
                                 "#FC4E07", "#b20a2c"))
set.seed(123)
p <- Heatmap(plotdata, 
             col = col_fun, column_split = clinicaldata$Subtypes,
             top_annotation = sub,
             show_row_names = F,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = F,
             cluster_columns = F, 
             cluster_rows = F, 
             row_names_gp = gpar(fontsize = 6),
             heatmap_legend_param = list(title = "Zscores"))
```

####  Neutrophil degranulation corrplot
```{r}
library(dplyr)
library(ggplot2)
library(ggpubr)
data <- openxlsx::read.xlsx("./NMF_C3_neu-pufa.xlsx", sheet = 1, rowNames = T)%>%as.data.frame()

plotdata <- as.data.frame(scale(log2(data)))

p <- ggplot(plotdata, aes(y=ALOX5, x=MMP9)) +
  geom_point(color="black") + stat_smooth(method="lm")+
  stat_cor(data=plotdata, method = "pearson")

m <- ggplot(plotdata, aes(y=ALOX5, x=MPO)) +
  geom_point(color="black") + stat_smooth(method="lm")+
  stat_cor(data=plotdata, method = "pearson")

i <- ggplot(plotdata, aes(y=ALOX5, x=ITGAM)) +
  geom_point(color="black") + stat_smooth(method="lm")+
  stat_cor(data=plotdata, method = "pearson")

g <- ggarrange(ggarrange(p, m, i,
                   labels = c("", "", ""),
                   ncol = 3, nrow = 1))

aa <- ggplot(plotdata, aes(y=ALOX5, x=Arachidonic.acid)) +
  geom_point(color="black") + stat_smooth(method="lm")+
  stat_cor(data=plotdata, method = "pearson")

la <- ggplot(plotdata, aes(y=ALOX5, x=Linoleic.acid)) +
  geom_point(color="black") + stat_smooth(method="lm")+
  stat_cor(data=plotdata, method = "pearson")
```

#### PUFA metabolism
```{r}
library(dplyr)
library(tibble)
library(circlize)
library(ComplexHeatmap)
pufa.data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolome_PUFA.xlsx", sheet = 1, rowNames = T)%>%as.data.frame()

pufa.data <- as.data.frame(t(pufa.data[,-c(1,13:19)]))

plotdata <- as.data.frame(t(scale(log2(1+t(pufa.data)))))

clinicaldata <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 3, rowNames = T)%>%as.data.frame()

sub = HeatmapAnnotation(Subtypes = clinicaldata$Subtypes,
                        show_legend = T,annotation_name_side = "left",
            col = list(Subtypes = c("NAT"= "grey",
                                    "C1" = "#2874C5",
                                    "C2" = "#EABF00",
                                    "C3" = "#C6524A")))

col_fun = circlize::colorRamp2(c(-4,-2, 0, 2, 4),
                               c("#49a09d","#00AFBB","white",
                                 "#FC4E07", "#b20a2c"))
set.seed(111)
p <- Heatmap(plotdata,
             col = col_fun,column_split = clinicaldata$Subtypes,
             top_annotation = sub,
             show_row_names = T,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = F,
             cluster_columns = F, 
             cluster_rows = F, 
             row_names_gp = gpar(fontsize = 7.5),
             heatmap_legend_param = list(title = "Zscores"))

```

#### Left and Right lobes
```{r}
library(dplyr)
library(tibble)
library(circlize)
library(ComplexHeatmap)
pro.list <- openxlsx::read.xlsx("./2022proteome/Proteome_hubgene_RP and RBs", sheet = 1, rowNames = F)%>%as.data.frame()

prodata <- openxlsx::read.xlsx("./2022proteome/2022Proteome.xlsx", sheet = 1, rowNames = F)%>%as.data.frame()

data <- merge(pro.list, prodata, by="Name",sort = F)
data <- column_to_rownames(data, var = "Name")

plotdata <- as.data.frame(t(scale(log2(1+t(data)))))

clinicaldata <- openxlsx::read.xlsx("./2022proteome/Clinicaldata2.xlsx", sheet = 2, rowNames = T)%>%as.data.frame()

col_lianxu <- colorRamp2(c(0, 10,40),  c("#f2fcfe","#6dd5fa","#2980b9"))

ht = HeatmapAnnotation(TBIL    = clinicaldata$TBIL,
                       show_legend  = T, annotation_name_side = "left",
                       col = list(TBIL=col_lianxu))

sub = HeatmapAnnotation(Subtypes = clinicaldata$Subtypes,
                        show_legend = T,annotation_name_side = "left",
            col = list(Subtypes = c("NAT"= "grey",
                                    "C1" = "#2874C5",
                                    "C2" = "#EABF00",
                                    "C3" = "#C6524A")))
seg = HeatmapAnnotation(Segments = clinicaldata$Segments,
                        show_legend = T,annotation_name_side = "left",
             col = list(Segments  = c("S1" = "#48466d", "S2"   = "#3d84a8",
                                     "S3" = "#46cdcf", "S4"   = "#abedd8",
                                     "S5" = "#f9ed69", "S6"   = "#f08a5d",
                                     "S7" = "#b83b5e", "S8"   = "#6a2c70")))
hlob = HeatmapAnnotation(Hepatic_lobes = clinicaldata$Hepatic_lobes,
                        show_legend = T,annotation_name_side = "left",
             col = list(Hepatic_lobes=c("NAT"= "grey",
                                        "Left_lobe"= "#0f4392",
                                        "Right_lobe"= "#ff5151")))

col_fun = circlize::colorRamp2(c(-4,-2, 0, 2, 4),
                               c("#49a09d","#00AFBB","white",
                                 "#FC4E07", "#b20a2c"))
set.seed(111)
p <- Heatmap(plotdata,
             #column_dend_reorder = order(clinicaldata$TBIL, decreasing = TRUE),
             col = col_fun,column_split = clinicaldata$Hepatic_lobes,
             top_annotation = c(hlob,ht,sub,seg),
             show_row_names = T,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = F,
             cluster_columns = T, 
             cluster_rows = F, 
             row_names_gp = gpar(fontsize = 7.5),
             heatmap_legend_param = list(title = "Zscores"))

```

#### Clinical information
```{r}
library(ggplot2)
library(ggpubr)
clinicaldata <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx",sheet = 2, rowNames = T)
##
clinicaldata$`ALB/GLB` <- factor(clinicaldata$Subtypes)

kruskal.test(`AST/ALT` ~ Subtypes, data = clinicaldata)
aov1 <- aov(BMI ~ Subtypes, data = clinicaldata)

my_comparisons <- list( c("C1", "C2"),c("C1", "C3"),
                        c("C2", "C3"))
p <- ggplot(clinicaldata, aes(x=Subtype, y=clinicaldata$`AST/ALT`, fill=Subtype))+
  geom_violin(alpha=0.5)+
  geom_boxplot(width=0.1,position = position_dodge(0.9))+
  stat_compare_means(comparisons = my_comparisons,
                     method = "t.test")+
  theme_classic()+ 
  theme(legend.position = "")+
  #ggtitle("BMI in Subtypes") + 
  #theme(plot.title = element_text(hjust = 0.5, size = 10))+
  xlab("") + ylab("AST/ALT")
```

### Metabolome&Lipidome NMF cluster
```{r}
library(NMF)
library(doMPI)
library(dplyr)
Lipid_data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Lipid2023.xlsx", sheet = 3, rowNames = T)%>%as.data.frame()

Lipid_data <- log2(1+Lipid_data)

## Metabolome
metabo.data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolites2.xlsx",sheet = 3, rowNames = T)%>%as.data.frame()

metabo.data <- log2(1+metabo.data)

res <- nmf(Lipid_data,2:8,nrun=10,seed=123)
plot(res)
## rank=3

res <- nmf(Lipid_data,3,nrun=10,seed=123)

group <- predict(res)
table(group)
write.csv(as.data.frame(group), file = "NMF-lipid_subtypes3.csv")

coefmap(res)
consensusmap(res)

index <- extractFeatures(res,0.95) 
sig.order <- unlist(index)
NMF.Exp <- Lipid_data[sig.order,]
NMF.Exp <- na.omit(NMF.Exp) 

library(tinyarray)
dp = NMF.Exp[,order(group)]
draw_heatmap(dp,sort(group), split_column = 3,
             color_an = c("#2874C5","#EABF00","#C6524A","#868686"),
             annotation_legend = T,
             cluster_cols = F,
             show_rownames = T)


jco <- c("#2874C5","#EABF00","#C6524A")
res2 <- nmf(NMF.Exp,3,nrun=10,seed=123)

coefmap(res2)
consensusmap(res2)
group2 <- predict(res2)

basismap(res, distfun="euclidean", hclustfun="ward",
 cexCol = 1,
 cexRow = 0.3,
 annColors=list(c("1"=jco[1],"2"=jco[2],"3"=jco[3])))

group <- predict(res) 
table(group)
consensusmap(res,
             labRow = NA,
             labCol = NA,
             annCol = data.frame("cluster"=group[colnames(res)]),
             annColors = list(cluster=c("1"=jco[1],"2"=jco[2],"3"=jco[3])))

```

#### Lipid and Metabolites
```{r}
library(tibble)
library(dplyr)
library(ComplexHeatmap)
library(circlize)
nmf_list <- openxlsx::read.xlsx("./NMF_C1-2-3_DALs.xlsx", sheet = 2, rowNames = F)
nmf_list <- as.data.frame(nmf_list[,1])
names(nmf_list) <- "Name" 

nmf_list <- nmf_list%>% distinct(Name, .keep_all = T)

metabo.data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolites2.xlsx", sheet = 4, rowNames = F)%>%as.data.frame()

dep.data <- merge(nmf_list, metabo.data, by = "Name", sort = FALSE)
dep.data <- column_to_rownames(dep.data, var = "Name")

plotdata <- as.data.frame(t(scale(log2(t(dep.data)))))

clinicaldata <- openxlsx::read.xlsx("./2022MBR/Clinicaldata.xlsx", sheet = 3, rowNames = T)%>%as.data.frame()

clinicaldata <- as.data.frame(clinicaldata[1:39,])

pc = HeatmapAnnotation(Subtypes = clinicaldata$Subtypes,
                        show_legend = T,annotation_name_side = "left",
            col = list(Subtypes = c("C1" = "#2874C5",
                                    "C2" = "#EABF00",
                                    "C3" = "#C6524A")))
lc = HeatmapAnnotation(LC.subtypes=clinicaldata$LC_subtypes,
                        show_legend = T,annotation_name_side = "left",
            col = list(LC.subtypes=c("LC1" = "#EABF00",
                                    "LC2" = "#2874C5",
                                    "LC3" = "#C6524A")))
mc = HeatmapAnnotation(MC.subtypes=clinicaldata$MC_subtypes,
                        show_legend = T,annotation_name_side = "left",
            col = list(MC.subtypes=c("MC1" = "#2874C5",
                                    "MC2" = "#C6524A",
                                    "MC3" = "#EABF00")))


col_fun = circlize::colorRamp2(c(-4,-2, 0,2, 4),
                               c("#49a09d","#00AFBB","white",
                                 "#FC4E07", "#b20a2c"))
set.seed(111)
p <- Heatmap(plotdata,
             col = col_fun, column_split = clinicaldata$Subtypes,
             top_annotation = c(pc,lc,mc),
             show_row_names = F,
             show_row_dend = F,
             show_column_dend = F,
             show_column_names = F,
             cluster_columns = F, 
             cluster_rows = F, 
             row_names_gp = gpar(fontsize = 6),
             heatmap_legend_param = list(title = "Relative Abundance"))


kruskal.test(PC_subtypes ~ Subtypes, data = clinicaldata)

kruskal.test(MC_subtypes ~ Subtypes, data = clinicaldata)

```

#### Metaboanalysis anotation
```{r}
library(ggplot2)
library(dplyr)
plotdata <- read.csv("./2022ProteomeMetabolism/NMF_C3_pathway_results.csv", header = T)
plotdata <- plotdata[1:6,]

p <- ggplot(plotdata, aes(y = reorder(factor(Name),logP),x=logP)) + 
     geom_bar(stat = 'identity', position = 'dodge', fill = "#C6524A") +
     theme_classic()+scale_x_reverse()+
     scale_y_discrete(position = "right")+
     xlab("-log10 (p-value)")+ylab("")

```

##CPTAC data

#### NMF subtypes in CPTAC
```{r}
library(dplyr)
library(tibble)
nmf.list <- read.csv("./NMF_log2_mad4000.csv", header = T)
nmf.list <- as.data.frame(nmf.list[,1])
names(nmf.list) <- "Symbol" 

nmf.list <- nmf.list%>% distinct(Symbol, .keep_all = T)

cptac.data <- openxlsx::read.xlsx("./Download data/hcc_prodata_tvsn.xlsx", sheet = 1, rowNames = F)

nmf.data <- merge(nmf.list, cptac.data, by="Symbol")
nmf.data <- column_to_rownames(nmf.data, var = "Symbol")

```

>NMF cluster

```{r}
library(NMF)
library(doMPI)
library(dplyr)

pro.exp <- log2(1+nmf.data)

res <- nmf(pro.exp,2:8,nrun=10,seed=123)
plot(res)
## rank=3

res <- nmf(pro.exp,3,nrun=10,seed=123)

group <- predict(res)
table(group)
write.csv(as.data.frame(group), file = "CPTAC_subtypes3.csv")

coefmap(res)
consensusmap(res)

jco <- c("#2874C5","#EABF00","#C6524A")

basismap(res, distfun="euclidean", hclustfun="ward",
 cexCol = 1,
 cexRow = 0.3,
 annColors=list(c("1"=jco[1],"2"=jco[2],"3"=jco[3])))

consensusmap(res,
             labRow = NA,
             labCol = NA,
             annCol = data.frame("cluster"=group[colnames(res)]),
             annColors = list(cluster=c("1"=jco[1],"2"=jco[2],"3"=jco[3])))

```

####CPTAC Ribosome-genes
```{r}
library(survminer)
library(survival)
library(dplyr)
library(tibble)
rb.list <- openxlsx::read.xlsx("./NMF_Ribosome b_genes2.xlsx", sheet = 2,rowNames = F)%>%as.data.frame()
rb.list <- as.data.frame(rb.list[,1])
names(rb.list) <- "Symbol" 

rb.list <- rb.list%>% distinct(Symbol, .keep_all = T)

cptac.data <- openxlsx::read.xlsx("./Download data/hcc_proteome_TTexp.xlsx", sheet = 1, rowNames = F)

rb.data <- merge(rb.list, cptac.data, by="Symbol")
rb.data <- as.data.frame(t(column_to_rownames(rb.data, var = "Symbol")))
rb.data <- rownames_to_column(rb.data, var = "TumorID")

CPTAC.clinical <- openxlsx::read.xlsx("./Download data/Zhou_Liver_Cancer_ProtemoeMeta.xlsx", sheet = 2, rowNames = F)

surv.data <- merge(CPTAC.clinical[,c(1,24:27)], rb.data, by="TumorID")

cutoff <- surv_cutpoint(surv.data, time = 'Survival_time', 
                        event = 'Survival_test', 
                        variables = c("XPO1","RPRD1B","RAN","CSNK2A1"))

summary(cutoff)

ncut <- cut(surv.data$XPO1, breaks = c(0, 1.798, Inf), 
            labels = c('Low', 'High'))
scut <- cut(surv.data$RPRD1B, breaks = c(0, 1.372, Inf), 
            labels = c('Low', 'High'))
pcut <- cut(surv.data$RAN, breaks = c(0, 2.185, Inf), 
            labels = c('Low', 'High'))
ccut <- cut(surv.data$CSNK2A1, breaks = c(0, 1.68, Inf), 
            labels = c('Low', 'High'))

fit <- survfit(Surv(Survival_time, Survival_test)~scut, data = surv.data)

p <- ggsurvplot(fit, pval=TRUE, surv.median.line = "hv", 
           palette=c("blue", "red"),
           legend.title="RPRD1B", legend.lab=c('Low','High'),
           legend=c(0.85,0.15),
           ggtheme = theme_bw(),
           title="Overall Survival in CPTAC")
p

```

####CPTAC ND-genes
```{r}
library(survminer)
library(survival)
library(dplyr)
library(tibble)
ne.list <- openxlsx::read.xlsx("./NMF_C1-2-3_DEPs.xlsx", sheet = 3,rowNames = F)
ne.list <- as.data.frame(ne.list[,1])
names(ne.list) <- "Symbol" 

ne.list <- ne.list%>% distinct(Symbol, .keep_all = T)

cptac.data <- openxlsx::read.xlsx("./Download data/hcc_proteome_TTexp.xlsx", sheet = 1, rowNames = F)

ne.data <- merge(ne.list, cptac.data, by="Symbol")
ne.data <- as.data.frame(t(column_to_rownames(ne.data, var = "Symbol")))
ne.data <- rownames_to_column(ne.data, var = "TumorID")

CPTAC.clinical <- openxlsx::read.xlsx("./Download data/Zhou_Liver_Cancer_ProtemoeMeta.xlsx", sheet = 2, rowNames = F)

surv.data <- merge(CPTAC.clinical[,c(1,24:27)], ne.data, by="TumorID")

cutoff <- surv_cutpoint(surv.data, time = 'Survival_time', 
                        event = 'Survival_test', 
                        variables = c("ICAM1","CEACAM8","ITGAM","S100A8",
                                      "ITGB2","LCN2","MPO","TIMP1","VCAM1",
                                      "MMP9","CD63","S100A9","ALOX5"))

summary(cutoff)

mcut <- cut(surv.data$MMP9, breaks = c(0, 1.900, Inf), 
            labels = c('Low', 'High'))
ccut <- cut(surv.data$CD63, breaks = c(0, 3.652, Inf), 
            labels = c('Low', 'High'))
scut <- cut(surv.data$S100A9, breaks = c(0, 0.222, Inf), 
            labels = c('Low', 'High'))
acut <- cut(surv.data$ALOX5, breaks = c(0, 5.196, Inf), 
            labels = c('Low', 'High'))

cut1 <- cut(surv.data$ICAM1, breaks = c(0, 0.938, Inf), 
            labels = c('Low', 'High'))
cut2 <- cut(surv.data$CEACAM8, breaks = c(0, 2.165, Inf), 
            labels = c('Low', 'High'))
cut3 <- cut(surv.data$ITGAM, breaks = c(0, 6.538, Inf), 
            labels = c('Low', 'High'))
cut4 <- cut(surv.data$S100A8, breaks = c(0, 0.439, Inf), 
            labels = c('Low', 'High'))
cut5 <- cut(surv.data$ITGB2, breaks = c(0, 2.480, Inf), 
            labels = c('Low', 'High'))
cut6 <- cut(surv.data$LCN2, breaks = c(0, 0.751, Inf), 
            labels = c('Low', 'High'))
cut7 <- cut(surv.data$MPO, breaks = c(0, 2.626, Inf), 
            labels = c('Low', 'High'))
cut8 <- cut(surv.data$TIMP1, breaks = c(0, 14.636, Inf), 
            labels = c('Low', 'High'))
cut9 <- cut(surv.data$VCAM1, breaks = c(0, 0.097, Inf), 
            labels = c('Low', 'High'))

fit <- survfit(Surv(Survival_time, Survival_test)~cut9, data = surv.data)

p <- ggsurvplot(fit, pval=TRUE, surv.median.line = "hv", 
           palette=c("blue", "red"),
           legend.title="VCAM1", legend.lab=c('Low','High'),
           legend=c(0.85,0.15),
           title="Overall Survival in CPTAC")

a <- ggsurvplot(fit, pval=TRUE, surv.median.line = "hv", 
           palette=c("blue", "red"),
           legend.title="ALOX5", legend.lab=c('Low','High'),
           legend=c(0.85,0.15),
           ggtheme = theme_bw(),
           title="Overall Survival in CPTAC")

```


#### HSF1 survplot
```{r}
library(survminer)
library(survival)
library(dplyr)
library(tibble)
rb.list <- openxlsx::read.xlsx("./NMF_C1-2-3_DEPs.xlsx", sheet = 4,rowNames = F)
rb.list <- as.data.frame(rb.list[,1])
names(rb.list) <- "Symbol" 
##name
rb.list <- rb.list%>% distinct(Symbol, .keep_all = T)

cptac.data <- openxlsx::read.xlsx("./Download data/hcc_prodata_tvsn.xlsx", sheet = 1, rowNames = F)

rb.data <- merge(rb.list, cptac.data, by="Symbol")
rb.data <- as.data.frame(t(column_to_rownames(rb.data, var = "Symbol")))
rb.data <- rownames_to_column(rb.data, var = "TumorID")

CPTAC.clinical <- openxlsx::read.xlsx("./Download data/Zhou_Liver_Cancer_ProtemoeMeta.xlsx", sheet = 2, rowNames = F)

surv.data <- merge(CPTAC.clinical[,c(1,24:27)], rb.data, by="TumorID")

cutoff <- surv_cutpoint(surv.data, time = 'Relapse_time', 
                        event = 'DFS_test', 
                        variables = c("HSF1"))

summary(cutoff)

mcut <- cut(surv.data$HSF1, breaks = c(0, 6.63, Inf), 
            labels = c('Low', 'High'))
fit <- survfit(Surv(Relapse_time, DFS_test)~mcut, data = surv.data)

p <- ggsurvplot(fit, pval=TRUE, surv.median.line = "hv", 
           palette=c("blue", "red"),
           legend.title="HSF1", legend.lab=c('Low','High'),
           title="KM Curve for DFS in CPTAC")
```

#### PUFA
```{r}
library(ggplot2)
library(ggpubr)
library(reshape2)
fa.data <- openxlsx::read.xlsx("./2022ProteomeMetabolism/FA.xlsx", sheet = 1, rowNames = T)
group <- openxlsx::read.xlsx("./2022ProteomeMetabolism/Metabolites.xlsx", sheet = 7, rowNames = T)

plotdata <- log2(fa.data+1)
plotdata <- cbind(group[[2]], plotdata)
names(plotdata) <- c("Group", "FA(20:4)","FA(20:5)", "FA(22:6)")

plotdata <- melt(plotdata, variable.name = "FA")

p <- ggplot(plotdata, aes(x=FA, y=value, fill=Group))+
  geom_violin(trim = FALSE, color="white", alpha=0.5, scale = "width")+
  geom_boxplot(width=0.1,position = position_dodge(0.9))+
  scale_fill_manual(values=c("#085f63","#49beb7","#fccf4d","#ef255f", "grey60"))+
  theme_bw()+ ylab("Relative Abundance")+xlab("")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

